### 2017-12-08 ABB_Fish.Rmd
updating files

---
title: "Antigua-Barbuda Fish Community Data: August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    df_print: paged
---
# Fish community data 
*required files: fish.csv*

## Part 1. Setup

### 1.1 Load packages
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(lazyeval)
library(rmarkdown)
library(tidyr)
library(ggplot2)
library(grid)
library(dplyr)
```

### 1.2 Import and format Antigua-Barbuda and Bonaire files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
AB <-read.csv("fish.csv", stringsAsFactors=FALSE)
AB <- AB[rep(row.names(AB), AB$Number), 1:18] # expand to replicate rows if multiple fish were recorded
AB <- AB %>% select(-c(Time,Depth.ft,Surveyor,Number,Date))

bonherb<-read.csv("Bonaire_herbivores.csv", stringsAsFactors=FALSE) %>% filter(Date==2017) %>% select(-Date) 
boncarn<-read.csv("Bonaire_carnivores_2017.csv", stringsAsFactors=FALSE)
boncarn <- boncarn[rep(row.names(boncarn), boncarn$Number), 1:16] %>% select(-c(Number,Diver,Date))
bonaire <- bind_rows(bonherb,boncarn) %>% filter(Site=="Oil Slick" | Site=="Karpata" | Site=="Front Porch" | Site=="18th Palm")

# bind Bonaire and AB data
fish <- bind_rows(AB,bonaire)

fish[fish==""]  <- NA 
fish[fish=="#NUM!"] <- NA
fish$Biomass <- as.numeric(fish$Biomass)
denuder <- fish %>% filter(Family=="Scaridae" | Family=="Acanthuridae")
scarid <- fish %>% filter(Family=="Scaridae")
acan <- fish %>% filter(Family=="Acanthuridae")
carnivore <- fish %>% filter(Family=="Serranidae" | Family=="Lutjanidae" | Family=="Carangidae") # need to expand appropriately, could specify potential predators
```

## Part 2. Summary Data
*2.1 Total fish populations, by site
*2.2 Total fish populations, by species and site
*2.2 Total denuder populations, by site
*2.3 Total scarid populations, by site
*2.4 Scarid populations, by species and site
*2.5 Scarid populations, by phase and site
*2.6 Total carnivore populations, by site

### 2.1 Total fish population, by site
```{r,message=FALSE,warning=FALSE}
totfish.tran<- fish %>% filter(Biomass != 0) %>% group_by(Island,Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.totfish<-totfish.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), tot.BM=mean(BM.t), tot.BM.se=(sd(BM.t))/sqrt(n), tot.DEN=mean(DEN.t), tot.DEN.se=(sd(DEN.t))/sqrt(n), tot.L=mean(L.t), tot.L.se=(sd(L.t))/sqrt(n)) %>% select(-n) # average across transects within each site
boxplot(BM.t~Site, las=2, data = totfish.tran)
```

### 2.2 Total fish populations, by species and site

```{r,message=FALSE,warning=FALSE}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.species <- fish %>% expand(nesting(Species,Family),Site)
temp.species <- full_join(expand.species,expand.tran,by="Site")

# Then I can join species/transect data with this species/transect template, and replace empty rows with 0 values

# biomass and density, by site and species:
species.tran<- fish %>% filter(Biomass != 0) %>% group_by(Site,Transect,Species) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.species, by=c("Site","Transect","Species")) # join transect-level data with template
species.tran[is.na(species.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
species.site<-species.tran %>% group_by(Site,Species,Family) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
species.L.tran <- fish %>% group_by(Site,Transect,Species) %>% dplyr::summarize(L.t=mean(Length))
species.L.site<-species.L.tran %>% group_by(Site,Species) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
species.site <- species.site %>% left_join(species.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations
```

### 2.3 Total denuder populations, by site
```{r,message=FALSE,warning=FALSE}
herb.tran<- denuder %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.herb<-herb.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), herb.BM=mean(BM.t), herb.BM.se=(sd(BM.t))/sqrt(n), herb.DEN=mean(DEN.t), herb.DEN.se=(sd(DEN.t))/sqrt(n), herb.L.s=mean(L.t), herb.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.3 Total acanthurid populations, by site
```{r,message=FALSE,warning=FALSE}
acan.tran<- acan %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.acan<-acan.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), acan.BM=mean(BM.t), acan.BM.se=(sd(BM.t))/sqrt(n), acan.DEN=mean(DEN.t), acan.DEN.se=(sd(DEN.t))/sqrt(n), acan.L=mean(L.t), acan.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.4 Total scarid populations, by site
```{r,message=FALSE,warning=FALSE}
scarid.tran<- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2, L.t=mean(Length))
site.scar<-scarid.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), scar.BM=mean(BM.t), scar.BM.se=(sd(BM.t))/sqrt(n), scar.DEN=mean(DEN.t), scar.DEN.se=(sd(DEN.t))/sqrt(n), scar.L=mean(L.t), scar.L.se=(sd(L.t))/sqrt(n)) %>% select(-n)
```

### 2.5 Scarid populations, by species and site
```{r,message=FALSE,warning=FALSE}
scarid.species <- species.site %>% filter(Family=="Scaridae") %>% select(-Family,-n) %>% ungroup()

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
viride <- scarid.species %>% filter(Species=="Sparisoma viride") %>% select(-Species)
names(viride)[2:8] <- c("viride.BM","viride.BM.se","viride.DEN","viride.DEN.se","viride.L.n","viride.L","viride.L.se")
vetula <- scarid.species %>% filter(Species=="Scarus vetula") %>% select(-Species)
names(vetula)[2:8] <- c("vetula.BM","vetula.BM.se","vetula.DEN","vetula.DEN.se","vetula.L.n","vetula.L","vetula.L.se")
aurofren <- scarid.species %>% filter(Species=="Sparisoma aurofrenatum") %>% select(-Species)
names(aurofren)[2:8] <- c("aurofren.BM","aurofren.BM.se","aurofren.DEN","aurofren.DEN.se","aurofren.L.n","aurofren.L","aurofren.L.se")
site.scar.sp <- left_join(viride,vetula,by="Site") %>% left_join(aurofren,by="Site")
```

### 2.6 Scarid populations, by phase and site
I will expand again here to make template with all phase/site combinations, then join transect level data
```{r,message=FALSE,warning=FALSE}
expand.tran <- fish %>% group_by(Site,Transect) %>% dplyr::summarize()
expand.phase <- scarid %>% expand(Phase,Site)
temp.phase <- full_join(expand.phase,expand.tran,by="Site")

# scarid biomass and density, by site and phase:
phase.tran <- scarid %>% filter(Biomass != 0) %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(temp.phase, by=c("Site","Transect","Phase")) # join transect-level data with template
phase.tran[is.na(phase.tran)] <- 0 # convert NAs to 0 biomass/density in transects where a species is not present
phase.site<-phase.tran %>% group_by(Site,Phase) %>% dplyr::summarize(n=n(), BM=mean(BM.t), BM.se=(sd(BM.t))/sqrt(n), DEN=mean(DEN.t), DEN.se=(sd(DEN.t))/sqrt(n))

# length, by site and species (calculated separately because transects missing length values should not be replaced with 0 values)
phase.L.tran <- scarid %>% group_by(Site,Transect,Phase) %>% dplyr::summarize(L.t=mean(Length))
phase.L.site<-phase.L.tran %>% group_by(Site,Phase) %>% dplyr::summarize(L.n=n(),L=mean(L.t), L.se=(sd(L.t))/sqrt(L.n))
phase.site <- phase.site %>% left_join(phase.L.site) # sample size will be different for length (L.n - only transects where species was observed) vs. biomass/density (n - all transects) calculations

# spread to site-level df with columns for each species' variables (ran into errors with unite/spread, so instead will do separately for each species)
juvenile <- phase.site %>% filter(Phase=="j") %>% select(-Phase,-n)
names(juvenile)[2:8] <- c("juvenile.BM","juvenile.BM.se","juvenile.DEN","juvenile.DEN.se","juvenile.L.n","juvenile.L","juvenile.L.se")
initial <- phase.site %>% filter(Phase=="i") %>% select(-Phase,-n)
names(initial)[2:8] <- c("intial.BM","intial.BM.se","intial.DEN","intial.DEN.se","intial.L.n","intial.L","intial.L.se")
terminal <- phase.site %>% filter(Phase=="t") %>% select(-Phase,-n)
names(terminal)[2:8] <- c("terminal.BM","terminal.BM.se","terminal.DEN","terminal.DEN.se","terminal.L.n","terminal.L","terminal.L.se")
site.scar.ph <- left_join(juvenile,initial,by="Site") %>% left_join(terminal,by="Site")
```

### 2.7 Total carnivore populations, by site
I will need to use expanded template, because some transects had no serranid/lutjanids
*NEED TO ADD LENGTH HERE*
```{r,message=FALSE,warning=FALSE}
carnivore.tran<- carnivore %>% filter(Biomass != 0) %>% group_by(Site,Transect) %>% dplyr::summarize(BM.t=sum(Biomass)/1.2, DEN.t=n()/1.2) %>% right_join(expand.tran,by=c("Site","Transect"))
carnivore.tran[is.na(carnivore.tran)] <- 0 
site.carn<-carnivore.tran %>% group_by(Site) %>% dplyr::summarize(n=n(), carn.BM=mean(BM.t), carn.BM.se=(sd(BM.t))/sqrt(n), carn.DEN=mean(DEN.t), carn.DEN.se=(sd(DEN.t))/sqrt(n)) %>% select(-n)
```

## Part 3. Combine into one wide dataframe by site, export csv for integration with follow data
```{r,message=FALSE,warning=FALSE}
site.fish <- site.totfish %>% left_join(site.herb,by="Site") %>% left_join(site.acan,by="Site") %>% left_join(site.scar,by="Site") %>% left_join(site.scar.sp,by="Site") %>% left_join(site.scar.ph,by="Site") %>% left_join(site.carn,by="Site")

write.csv(site.fish,file="site.fish.csv")
write.csv(species.site,file="species.site.csv")
```