### 2017-12-08 ABB_Integration.Rmd

---
title: "Data integration"
output: html_document
---
# Integrating fish follow data with site-level fish, benthic, and rugosity data 
*Files: site.fish.csv, site.follows.csv, site.benthic.csv, site.human.csv, species.site.csv, sumID.csv*

## Part 1. Setup

```{r,message=FALSE,warning=FALSE}
# set up
setwd("~/Google Drive/Research + Projects/Dissertation/ABB_Grazing/Analysis")
library(lazyeval)
library(rmarkdown)
library(dplyr)
library(tidyr)
library(ggplot2)
library(grid)
```
### 1.2 Import and format files
```{r, echo=FALSE,message=FALSE,warning=FALSE}
site.follows<-read.csv("site.follows.csv",stringsAsFactors=FALSE)%>% select(-X)
site.fish<-read.csv("site.fish.csv", stringsAsFactors=FALSE)%>% select(-X)
site.benthic<-read.csv("site.benthic.csv", stringsAsFactors=FALSE)%>% select(-X)
site.human<-read.csv("site.human.csv", stringsAsFactors=FALSE)
species.site<-read.csv("species.site.csv",stringsAsFactors = FALSE)%>% filter(Species=="Sparisoma viride" | Species=="Scarus vetula" | Species=="Sparisoma aurofrenatum") %>% select(-X,-Family,-n,-L.n)
names(species.site)[3:8] <- c("conspec.BM","conspec.BM.se","conspec.DEN","conspec.DEN.se","conspec.L","conspec.L.se")
sumID <- read.csv("sumID.csv", stringsAsFactors=FALSE)%>% select(-X)
```
### 1.3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
site.sum <- site.follows %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site") 
# then join site-level data to sumID dataframe
sumID2 <- sumID %>% left_join(site.fish,by="Site") %>% left_join(site.benthic,by="Site") %>% left_join(site.human,by="Site")
# calculate conspecifc scarid variable to add to sumID2 (based on fish site/species)
sumID2 <- sumID2 %>% left_join(species.site,by=c("Site","Species"))

write.csv(site.sum,file="site.sum.csv")
write.csv(sumID2,file="sumID2.csv")
```