### 2018-01-31 Models.Rmd
update 1/31

---
title: "Model Trials"
author: "Molly Wilson"
date: "1/6/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---
```{r}
library(rmarkdown)
library(dplyr)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
```

```{r}
setwd("~/github/Scarid-Behavior/data")
data <- read.csv("data2.csv", stringsAsFactors=FALSE)
vir<-data%>%filter(Species.Code=="stop")
vet<-data%>%filter(Species.Code=="qup")
auro<-data%>%filter(Species.Code=="rbp")
vetvir<-data%>%filter(Species.Code!="rbp")
```

# Model trials

## Basic lm
-site-level predictors: scar.BM, scar.DEN, benthic (PC1,PC2)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar.BM interaction?

```{r}
lm_FR <- lm(FR~scar.BM*Species+Phase+Length.cm+PC1+PC2, data=data)
summary(lm_FR)
```
*Notes: nearly everything is significant? Significant interaction between scar.BM and species for Spar. viride.*
*R2 goes up slightly with initial only*

### Adding spearfishing variable
```{r}
lm_FR <- lm(FR~scar.BM*Species+Phase+Length.cm+PC1+PC2+Spearfishing, data=data)
summary(lm_FR)
```
*Notes: spearfishing masks benthic (PC1/PC2 effects), because highly correlated...*

## Mixed effects models
- random effect: island

```{r}
lme_FR <- lme(FR~Phase+Length.cm+Species*scar.BM+PC1+PC2+carn.BM,data=data,random=~1|Island, na.action = na.omit)
summary(lme_FR)
```
*Notes: including island as a random effect makes all site-level variables non-significant. Adding carn.BM back in makes *

### Fraction of time spent grazing
```{r}
lme_g.frac <- lme(g.frac~Phase+Length.cm+Species*scar.BM+scar.DEN+PC1+PC2,data=data,random=~1|Island, na.action = na.omit)
summary(lme_g.frac)
```
*Notes: carn.BM and phase are strongest predictors, followed by scar.BM, PC1, and length. Also significant interaction between Spar. viride and scar.BM*

## GAMs
```{r, echo=FALSE}
gam<-gam(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, method="REML")
summary(gam)
plot(gam,pages=1)
```
```{r, echo=FALSE}
gamvet<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+(carn.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vet, method="REML")
summary(gamvet)
plot(gamvet,pages=1)
```
```{r, echo=FALSE}
vir<-data %>% filter(Species.Code=="stop" & Site!="Snapper Point")
gamvir<-gam(FR~Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vir,method="REML")
summary(gamvir)
plot(gamvir,pages=1)
```

## GAMMs
```{r, echo=FALSE}
gamm<-gamm(FR~Species+Phase+s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=data, random=list(Island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```