### 2018-04-04 SummaryOrals.Rmd
4/4

---
title: 'Summary: Dissertation Proposal'
author: "Molly Wilson"
date: "3/14/2018"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
    toc_depth: 3
  pdf_document:
    toc: yes
---

This is a summary of the primary figures and tests I have completed for Chapter 2 as of my oral exams (April 2018), as well as accompanying notes and questions.

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(rmarkdown)
library(dplyr)
library(tidyr)
library(lazyeval)
library(ggplot2)
library(ellipse)
library(RColorBrewer)
library(ggbiplot)
library(nlme)
library(pls)
library(mgcv)
library(ggsignif)
library(car)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
setwd("~/github/Scarid-Behavior/data")
site.sum <- read.csv("site.sum.csv", stringsAsFactors=FALSE) %>% select(-X)
data <- read.csv("data2.csv", stringsAsFactors=FALSE)

vetvir <- data %>% filter(Species.Code!="rbp")
vetvir20 <- vetvir %>% filter(Length.cm>=15 & Length.cm<=25)
init20<- vetvir %>% filter(Length.cm>=15 & Length.cm<=25 & Phase=="i")
vetvirAB <- vetvir %>% filter(Island!="Barbuda")
vetvirAB20 <- vetvir %>% filter(Island!="Barbuda")
vetvirABinit <- vetvirAB %>% filter(Phase=="i")

vetvirAB20NS <- vetvirAB20 %>% filter(Site!="Snapper Point")
vetvir20NS <- vetvir20 %>% filter(Site!="Snapper Point")
vetvirNS <- vetvir %>% filter(Site!="Snapper Point")
```

# Correlations among predictor variables

## Benthic variables
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~cover_CCA+cover_LC+cover_MA+canopy_MA+cover_TA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Benthic Community Variables",na.action = na.omit)
```

## Selected fish and benthic variables

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar.BM+scar.DEN+carn.BM+cover_LC+cover_TA+canopy_MA+canopy_TA+rugosity, data=data,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

# PCAs

## All predictors
```{r}
predictors <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity,scar.BM,scar.DEN) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE) 
summary(pca)

plot(pca,type="l")
ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()
ggsave("pca_pred.png")
```

## Benthic predictors

```{r, echo=FALSE}
pred.benthic <- site.sum %>% select(cover_LC,cover_TA,canopy_MA,canopy_TA,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?)
pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)

ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=site.sum$Island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="Island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal()

# add PC1 and PC2 as variables in dataset
pca1.scores <- as.data.frame(pca1$x) %>% select(PC1,PC2)
site.sum <- site.sum  %>% bind_cols(pca1.scores)
pca1.scores <- select(site.sum,Site,PC1,PC2)
data2 <- data %>% left_join(pca1.scores,by="Site")

setwd("~/github/Scarid-Behavior/data")
write.csv(data2,file = "data2.csv")
```

# Grazing differences between islands

## Feeding rates

```{r, message=FALSE,warning=FALSE,echo=FALSE}
vetvir20init<-vetvir20 %>% filter(Phase=="i")
ggplot(vetvir20init, aes(x=Island, y=FR)) + geom_boxplot() +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Feeding Rate (bites/hr)",x="",title="Feeding Rate by Island and Species (Initial Phase)") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
ggsave("bp_fr.png")
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
vet20init<- vetvir20init %>% filter(Species.Code=="qup")
aovFRvet <- aov(FR ~ Island, data=vet20init) 
summary.aov(aovFRvet) # significant differences
TukeyHSD(aovFRvet)
plot(aovFRvet, 2) # normally distributed
leveneTest(FR ~ Island, data = vet20init) # homogeneity of variance

vir20init<- vetvir20init %>% filter(Species.Code=="stop")
aovFRvir <- aov(FR ~ Island, data=vir20init, white.adjust=TRUE) 
summary.aov(aovFRvir) # significant differences
TukeyHSD(aovFRvir)
plot(aovFRvir, 2) # normally distributed
leveneTest(FR ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
```

## Bite rates

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init, aes(x=Island, y=BR)) + geom_boxplot() +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Bite Rate (bites/sec)",x="",title="Bite Rate while Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
ggsave("bp_br.png")
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
aovBRvet <- aov(BR ~ Island, data=vet20init, white.adjust=TRUE) 
summary.aov(aovBRvet) # no significant differences
TukeyHSD(aovBRvet)
plot(aovBRvet, 2) # fairly normal?
leveneTest(BR ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE

aovBRvir <- aov(BR ~ Island, data=vir20init, white.adjust=TRUE) 
summary.aov(aovBRvir) # no significant differences
TukeyHSD(aovBRvet)
plot(aovBRvir, 2) # fairly normal?
leveneTest(BR ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
```

## Fraction of time spent grazing

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init, aes(x=Island, y=g.frac)) + geom_boxplot() +
stat_summary(fun.y = mean, geom="point", shape=5, size=3) + labs(y="Time fraction",x="",title="Fraction of Time Spent Grazing by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
ggsave("bp_gfrac.png")
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
aovBRvet <- aov(BR ~ Island, data=vet20init, white.adjust=TRUE) 
summary.aov(aovBRvet) # no significant differences
plot(aovBRvet, 2) # fairly normal?
leveneTest(BR ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE

vir<- vetvir %>% filter(Species.Code=="stop")
aovBRvir <- aov(BR ~ Island, data=vir20init, white.adjust=TRUE) 
summary.aov(aovBRvir) # no significant differences
plot(aovBRvir, 2) # fairly normal?
leveneTest(BR ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
```

## Grazing intensity

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init, aes(x=Island, y=for.bites)) + geom_boxplot()  + labs(y="Bites per foray",x="",title="Grazing intensity by Island and Species") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) + facet_grid(.~Species)
ggsave("bp_gi.png")
```
```{r, message=FALSE,warning=FALSE,echo=FALSE}
aovGIvet <- aov(for.bites ~ Island, data=vet20init, white.adjust=TRUE) 
summary.aov(aovGIvet) # significant differences
TukeyHSD(aovGIvet)
plot(aovGIvet, 2) # fairly normally distributed?
leveneTest(for.bites ~ Island, data = vet20init) # heteroskedastic - add white.adjust=TRUE

aovGIvir <- aov(for.bites ~ Island, data=vir20init, white.adjust=TRUE) 
summary.aov(aovGIvir) # significant differences
TukeyHSD(aovGIvir)
plot(aovGIvir, 2) # fairly normally distributed?
leveneTest(for.bites ~ Island, data = vir20init) # heteroskedastic - add white.adjust=TRUE
```

# Bivariate plots

## Grazing as a function of fish size
*Due to the lack of larger fish in Barbuda, these length-based relationships focus on comparisons between Antigua and Bonaire*

```{r, message=FALSE,warning=FALSE,echo=FALSE}

ggplot(vetvirAB,aes(x=Length.cm,y=FR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Feeding rates by fish length") + theme_minimal() +labs(y="Feeding rates (bites/hr)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
ggsave("sp_length.png")
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirABinit,aes(x=Length.cm,y=BR,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Bite rate by fish length") + theme_minimal() +labs(y="Bite Rate (bites/sec)",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirABinit,aes(x=Length.cm,y=g.frac,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by fish length") + theme_minimal() +labs(y="Time fraction",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvirABinit,aes(x=Length.cm,y=for.bites,colour=Island))+facet_grid(.~Species)+geom_point(shape=18)+scale_shape(solid=FALSE)+geom_smooth(method=lm, fullrange=FALSE)+ggtitle("Grazing intensity by fish length") + theme_minimal() +labs(y="Bites/foray",x="Length (cm)") + theme(plot.title = element_text(hjust = 0.5))  + scale_color_manual(values=c("navy", "slategray2"))
```

## Grazing impact as a function of scarid biomass

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init,aes(x=scar.BM,y=FR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Feeding rate by scarid biomass") + theme_minimal() +labs(y="Feeding rate (bites/hr)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
ggsave("sp_scarbm.png")
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init,aes(x=scar.BM,y=BR))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Bite rate while grazing by scarid biomass") + theme_minimal() +labs(y="Bite rate (bites/sec)",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init,aes(x=scar.BM,y=g.frac))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Fraction of time spent grazing by scarid biomass") + theme_minimal() +labs(y="Time fraction",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(vetvir20init,aes(x=scar.BM,y=for.bites))+facet_grid(.~Species)+geom_point(shape=18)+geom_smooth(method=loess, fullrange=FALSE)+ggtitle("Grazing intensity by scarid biomass") + theme_minimal() +labs(y="Bites/foray",x="Scarid Biomass (g/100m2)") + theme(plot.title = element_text(hjust = 0.5))
```

# Competition

```{r,echo=FALSE,warning=FALSE}
ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_stop, shape=Island))+geom_point(size=3)+ggtitle("S. viride interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_stop-int.scar.se_stop, ymax=int.scar.mean_stop+int.scar.se_stop))
ggsave("comp_vir.png")
```

```{r,echo=FALSE,warning=FALSE}
ggplot(site.sum,aes(x=scar.BM,y=int.scar.mean_qup, shape=Island))+geom_point(size=3)+ggtitle("S. vetula interaction frequency as a function of scarid biomass") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) + labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + geom_errorbar(aes(ymin=int.scar.mean_qup-int.scar.se_qup, ymax=int.scar.mean_qup+int.scar.se_qup))
ggsave("comp_vet.png")
```

# GAMMs

## Scarus vetula

### Feeding rates
```{r,message=FALSE,warning=FALSE}
vetinit <- data2 %>% filter(Species.Code=="qup" & Phase=="i")
gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # only PC1 and PC2 are significant, length slightly, scar.BM is not
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # PC1 and PC2 are most significant, length is significant, scar.BM is almost significant
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```

### Grazing intensity
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=vetinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # very low R2, only PC1 and scar.BM significant
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```

## Sparisoma viride

### Feeding rates
```{r,message=FALSE,warning=FALSE}
virinit <- data2 %>% filter(Species.Code=="stop" & Phase=="i")
gamm<-gamm(FR~s(Length.cm)+s(scar.BM)+s(PC1)+s(PC2), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
summary(gamm$gam) #PC1, length, and scar.BM are significant predictors
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g.frac~s(Length.cm)+s(scar.BM)+s(PC1), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
summary(gamm$gam) #PC1, length, and scar.BM are significant predictors
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```

### Grazing intensity
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(for.bites~s(Length.cm)+s(scar.BM)+s(PC1), family=gaussian(link=identity),data=virinit, random=list(Island=~1),method="REML")
summary(gamm$gam) # no significant predictors (scar.BM almost), low r2
AIC(gamm$lme)
plot(gamm$gam,pages=1)
```