### 2018-11-13 r/Graphing+Analysis.Rmd
11/12

---
title: "Scarid grazing analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true

---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(startR)
library(here)
library(tidyverse)
# library(lazyeval)
# library(ellipse)
library(RColorBrewer)
#library(ggbiplot)
library(corrplot)
library(nlme)
library(pls)
library(factoextra)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")

sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  select(island,site,spearfishing,diving,species,species_code,phase,length_cm,g_frac,fr,br,tdb,for_dur,for_bites,tot_bm,tot_den,scar_bm,herb_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,cover_ta,canopy_ma,canopy_ta,rugosity,int_tot,int_scar,int_con,int_hetscar,int_dam)

stop<-sum_id%>%filter(species_code=="stop")
qup<-sum_id%>%filter(species_code=="qup")
rbp<-sum_id%>%filter(species_code=="rbp")
```

# Data overview:
- Fish follows (2 min) conducted at sites in Antigua (6), Barbuda (3) and Bonaire (4) from March-August 2017
- Follows tracked time spent grazing, bite rates, and competitive interactions (among scarids and with damselfish)
- Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
  +*Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species. Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
- site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites

```{r,echo=FALSE}
sample <- group_by(sum_id, site, species, phase) %>% summarize(n=n())
ggplot(data=sample, aes(x=site, y=n, fill=phase)) +
    geom_bar(stat="identity", position=position_dodge()) + facet_grid(.~species) + theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_grey()
# counts <- table(sum_id$species, sum_id$site)
# barplot(counts, main="Sample size by phase and species",
#   legend = rownames(counts), beside=TRUE, las=2)
```

# Examing variable distributions and relationships

## Site-level predictor variables: fish, benthic, and rugosity

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

### Correlations
```{r,echo=FALSE}
site_var <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,conspec_bm,conspec_den,carn_bm,carn_den,cover_cca,cover_lc,cover_ma,canopy_ma,cover_ta,canopy_ta,rugosity)
cor = cor(site_var, use="pairwise.complete.obs")
my_colors <- brewer.pal(5, "Spectral")
my_colors = colorRampPalette(my_colors)(100)
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

Many correlations here, both within benthic/fish categories and across categories. Rugosity is correlated with both fish and benthic variables. I will examine correlations within fish community and benthic community variables to decide which I can omit and which I will integrate through pcA. Whether I include benthic variables as predictor or response variables will determine if I need to worry about fish~benthic correlations.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot_bm+tot_den+scar_bm+scar_den+conspec_bm+conspec_den+carn_bm+carn_den+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables", na.action = na.omit)
```

Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric. Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them. Conspecific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove conspec_den?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm+scar_den+conspec_den+carn_bm+rugosity, data=sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspecific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~cover_cca+cover_lc+cover_ma+canopy_ma+cover_ta+canopy_ta+rugosity, data = sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main ="Benthic Community Variables", na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cover_cca and cover_ma (accounted for by canopy_ma). The rest should be combined via pcA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm+scar_den+carn_bm+cover_lc+cover_ta+canopy_ma+canopy_ta+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid_bm.s and all benthic variables....). It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables. Variables scar_den and conspec_den are relatively independent of all other variables (except each other).

Current outline for site-level pcA tasks:
If benthic=predictor:
- cover_lc + cover_ta + canopy_ma + canopy_ta + rugosity + scar_bm + carn_bm
- scar_den + conspec_den
If benthic=response:
- rugosity + scar_bm + carn_bm
- cover_lc + cover_ta + canopy_ma + canopy_ta
- scar_den + conspec_den

## Fish-level response variables: grazing behaviors

### Distributions
```{r,echo=FALSE,warning=FALSE}
grazing <- sum_id %>% 
  select(g_frac,fr,br,for_dur,for_bites) %>% 
  gather() %>% 
  na.omit()

ggplot(grazing,aes(value)) + 
  facet_wrap(~ key, scales = "free") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

### Corellations
```{r,echo=FALSE}
grazing <- sum_id %>% select(g_frac,fr,br,for_dur,for_bites)
cor=cor(grazing, use="pairwise.complete.obs")
ord <- order(cor[1, ])
cor_ord = cor[ord, ord]
plotcorr(cor_ord , col=my_colors[cor_ord*50+50] , mar=c(1,1,1,1)  )
```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

Variable for_bites is correlated with nearly all other variables, so will be omitted. Variables g.farc, fr. and for_dur are still correlated and should be assessed with a pcA. Variable br can be a standalone variable.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~g_frac+fr+br+for_dur, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Grazing Response Variables",na.action = na.omit)
```

Grazing pcA:
- g_frac + fr + for_dur
- (br as standalone)

# Exploratory bivariate plots

## Grazing as a function of fish length
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = br, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = g_frac, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Time Fraction", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```

## Grazing as a function of scarid biomass

```{r}
ggplot(sum_id %>% filter(island!="Barbuda"), aes(x = scar_bm,y=fr)) +
  facet_grid(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = island)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by phase and scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = br, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bite rates (while grazing) by scarid biomass") +
  labs(y = "Bite rate (bites/sec)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = g_frac, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Fraction of time spent grazing by phase and scarid biomass") +
  labs(y = "Time Fraction", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = for_bites, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bites per foray by phase and scarid biomass") +
  labs(y = "Bites", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```

## Competitive interactions

```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean, shape=island)) +
  geom_point(size=3) +
  ggtitle("Interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean-int_scar.se, ymax=int_scar.mean+int_scar.se))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_stop, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. viride interaction frequency as a function of scarid biomass") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Scarid Biomass (g/100m2)", y = "Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean_stop-int_scar.se_stop, ymax=int_scar.mean_stop+int_scar.se_stop))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_qup, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. vetula interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + 
  geom_errorbar(aes(ymin=int_scar.mean_qup-int_scar.se_qup, ymax=int_scar.mean_qup+int_scar.se_qup))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id,aes(x=scar_bm,y=int_scar)) +
  facet_grid(.~species)+geom_point(shape=19) +
  geom_smooth(method=lm) +
  ggtitle("Competitive interactions as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

<!-- # pcAs -->
<!-- *Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->

<!-- ## Predictors -->

<!-- ### All predictors -->
<!-- ```{r} -->
<!-- predictors <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity,scar_bm,scar_den,tot_bm) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca <- prcomp(na.omit(predictors),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca) -->

<!-- plot(pca,type="l") -->
<!-- ggbiplot(pca, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, varname.size = 2) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->
<!-- ggsave("pcApred.png") -->
<!-- ``` -->
<!-- ### Benthic variables -->
<!-- ```{r, echo=FALSE} -->
<!-- pred.benthic <- sum_site %>% select(cover_lc,cover_ta,canopy_ma,canopy_ta,rugosity) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca1 <- prcomp(na.omit(pred.benthic),center = TRUE,scale. = TRUE)  -->

<!-- ggbiplot(pca1, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE, varname.size=2.5, varname.adjust=1.1) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->

<!-- # add pc1 and pc2 as variables in dataset -->
<!-- pca1.scores <- as.data.frame(pca1$x) %>% select(pc1,pc2) -->
<!-- sum_site <- sum_site  %>% bind_cols(pca1.scores) -->
<!-- pca1.scores <- select(sum_site,site,pc1,pc2) -->
<!-- sum_id_pca <- sum_id %>% left_join(pca1.scores,by="site") -->

<!-- write.csv(sum_id_pca,file = here("data/sum_id_pca.csv")) -->
<!-- ``` -->
<!-- ### Fish community variables -->
<!-- ```{r} -->
<!-- pred.fish <- sum_site %>% select(scar_bm,scar_den,carn_bm) # should it matter if I use site level summary data vs. full follow dataset (where site-level data has been joined?) -->
<!-- pca2 <- prcomp(na.omit(pred.fish),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca2) -->

<!-- plot(pca2,type="l") -->
<!-- ggbiplot(pca2, obs.scale = 1, var.scale = 1, groups=sum_site$island, ellipse = TRUE, circle = TRUE, loadings.label.repel=TRUE) + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() -->

<!-- ``` -->

<!-- ## Response variables -->

<!-- ### Grazing behaviors -->
<!-- *log transform increases P1 explained variance here...* -->
<!-- *Currently for_dur has several NA values (for individuals that didn't graze)... I am omitting that variable for now because I don't want to omit non-grazing individuals (instead would like to keep them as grazing=0* -->
<!-- ```{r} -->
<!-- response <- data %>% select(g_frac,tdb,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca3) -->
<!-- plot(pca3,type="l") -->
<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1, groups=data$island,shape=data$island,ellipse = TRUE, circle = TRUE, alpha=0.8,varname.size = 5) + theme_minimal() + scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + labs(title="Grazing behaviors by island") + theme(plot.title = element_text(hjust = 0.5)) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- response <- data %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pca3 <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- summary(pca3) -->
<!-- plot(pca3,type="l") -->
<!-- ggbiplot(pca3, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=0.4,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() +geom_point(aes(col=data$scar_bm)) + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) -->
<!-- ``` -->

<!-- ```{r,echo=FALSE} -->
<!-- vir <- data %>% filter(species_code=="stop") -->
<!-- response <- vir %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pcavir <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- ggbiplot(pcavir, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vir$scar_bm)) -->
<!-- ``` -->
<!-- ```{r,echo=FALSE} -->
<!-- vet <- data %>% filter(species_code=="qup") -->
<!-- response <- vet %>% select(g_frac,fr,br) -->
<!-- log.vars5 <- log(response[,1:3]) -->
<!-- pcavet <- prcomp(na.omit(response),center = TRUE,scale. = TRUE)  -->
<!-- ggbiplot(pcavet, obs.scale = 1, var.scale = 1,ellipse = TRUE, circle = TRUE, alpha=1,varname.size = 5) + theme(legend.direction = 'horizontal', legend.position = 'top') + theme_minimal() + labs(col="Scarid Biomass (g/100m2)", title="Grazing Behaviors by Scarid Biomass") + theme(plot.title = element_text(hjust = 0.5)) +geom_point(aes(colour=vet$scar_bm)) -->
<!-- ``` -->

# Model trials

-site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar_bm interaction?
-random effects: island
-response variables: g_frac, br (?), and fr (run separately)

## Basic lm (to assess predictors of interest)
```{r}
initial<-data%>%filter(phase=="i") # for analysis with initial only
lm_fr <- lm(fr~scar_bm*species+phase+carn_bm+length_cm+pc1+pc2, data=data)
summary(lm_fr) # predicting feeding rate of initial phase only, all species
#vif(lm_fr)
```
*Notes: nearly everything is significant? Significant interaction between scar_bm and species for Spar. viride. carn_bm and scar_bm are highly positively correlated (~0.86) - is it ok to include both?*
*R2 goes up slightly with initial only*

```{r}
lm_g_frac <- lm(g_frac~species*scar_bm+phase+scar_den+carn_bm+phase+length_cm+pc1+pc2, data=data)
summary(lm_g_frac)
```
*Notes: carn_bm and pc1 are strongest predictors of time spent grazing, scar_bm and phase are also significant. Slight significance of species, length, and species-scar_bm interaction*

## Mixed effects models
- random effect: island

```{r}
lme_fr <- lme(fr~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
summary(lme_fr)
```
*Notes: nearly everything is significant except for pc2 and scar_den?*
```{r}
lme_g_frac <- lme(g_frac~phase+length_cm+species*scar_bm+scar_den+carn_bm+pc1+pc2,data=data,random=~1|island, na.action = na.omit)
summary(lme_g_frac)
```
```{r}
lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
summary(lme_fr_stop)
```
*Notes: carn_bm and length are signifigant predictors*
```{r}
lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
summary(lme_fr_qup)
```
*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*
# Model trials: species run individually

## Feeding rates
```{r}
lme_fr_stop <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=stop,random=~1|island, na.action = na.omit)
summary(lme_fr_stop)
```
*Notes: carn_bm and length are signifigant predictors*

```{r}
lme_fr_qup <- lme(fr~phase+length_cm+scar_bm+scar_den+carn_bm+pc1+pc2,data=qup,random=~1|island, na.action = na.omit)
summary(lme_fr_qup)
```
*Notes: pc1 is strongest (magnitude) predictors. Also significant are phase, carn_bm and scar_den*

## GAMs
```{r, echo=FALSE}
gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, method="REML")
summary(gam)
plot(gam,pages=1)
```
```{r, echo=FALSE}
gamvet<-gam(fr~phase+s(length_cm)+s(scar_bm)+(carn_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, method="REML")
summary(gamvet)
plot(gamvet,pages=1)
```
```{r, echo=FALSE}
vir<-data %>% filter(species_code=="stop" & site!="Snapper Point")
gamvir<-gam(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir,method="REML")
summary(gamvir)
plot(gamvir,pages=1)
```

## GAMMs

### Feeding rate (bites per hour)
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

Most of variation is explained by species - for individual species models R2 values are very low (~0.3)

### Fraction of time spent grazing
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=data, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vet, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(g_frac~phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=vir, random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

Low R2 for all trials predicting g_frac (<0.4)(gammvir)
```
------------------------
To Do as of Nov. 7
* ~~incorporate pcs~~
* boosted regression trees ecosphere 2017 adrians paper
* ~~simpler graphs with just herb density x axis - grazing behaviors y, facet by species~~
* ~~species*density as fixed effect~~
* species as random effect
* ~~break out species and run separately~~