### 2018-11-13 r/Analysis.Rmd
11/12 adjustments

---
title: "Figures + Analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Outline of selected figures and analyses:

- Grazing differences between islands: (site or island level?)
-- ANOVAs to test statistical significance of differences here
- Total grazing impact of site-level biomass: measured vs. reported grazing rates (add line or color by feeding rate)
- Something looking at relationship with algal index vs. scarid biomass  
- SEM?

Notes: 
- re: lms/gamms: don't focus on these for now, the data has so many colinearities... if anything, use full PCA in model analysis
- Exclude Barbuda from this analysis...?