### 2018-11-13 r/Data_prodding.Rmd
11/13

---
title: "Investigation of scarid data"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true

---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
library(startR)
library(here)
library(tidyverse)
#library(RColorBrewer)
library(corrplot)
library(ggplot2)

library(nlme)
library(gamm4)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Data overview:
- Fish follows (2 min) conducted at sites in Antigua (6), Barbuda (3) and Bonaire (4) from March-August 2017
- Follows tracked time spent grazing, bite rates, and competitive interactions (among scarids and with damselfish)
- Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
  +*Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species. Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
- site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites

```{r,echo=FALSE}
sample <- group_by(sum_id, site, species, phase, island) %>% summarize(n=n())

ggplot(sum_id, aes(site)) +
        facet_grid(island~species) +
        geom_bar(aes(fill = phase), position = "dodge") +
        scale_fill_manual(values = c("gray", "slategray2")) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))

# playing with plot options...
ggplot(data = sample, aes(x = site, y = n, fill = phase)) + 
    geom_bar(stat = "identity", width = 1, position = "dodge") +
    facet_wrap(~island, strip.position = "bottom", scales = "free_x") +
    theme(panel.spacing = unit(0, "lines"), 
         strip.background = element_blank(),
         strip.placement = "outside") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Examing variable distributions and relationships

## Site-level predictor variables: fish, benthic, and rugosity

### Distributions
```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn_bm,carn_den,cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```
Site-level predictors are not normally distributed....

### Correlations
```{r,echo=FALSE}
cor = cor(sum_id %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
         carn_bm,carn_den,cca_cover,lc_cover,
         ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity), 
  use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)

```

Many correlations here, both within benthic/fish categories and across categories. Rugosity is correlated with both fish and benthic variables. I will examine correlations within fish community and benthic community variables to decide which I can omit and which I will integrate through pcA. Whether I include benthic variables as predictor or response variables will determine if I need to worry about fish~benthic correlations.

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~tot_bm+tot_den+scar_bm+scar_den+consp_bm+consp_den+carn_bm+carn_den+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Fish Community Variables", na.action = na.omit)
```

Variables carn_den and carn_bm are essentially identical - I will remove carn_den and use carn_bm as the main preditor metric. Total fish biomass and density (tot_bm and tot_den) are highly correlated with both scarid and carnivore variables, so I will remove them. conspific biomass is highly correlated with scarid biomass so I will remove it as well (and should possibly also remove consp_den?). The selected variables I will proceed with are as follows:
```{r,echo=FALSE}
pairs(~scar_bm+scar_den+consp_den+carn_bm+rugosity, data=sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main = "Selected Fish Community Variables", na.action = na.omit)
```

Scarid biomass, carnivore biomass, and rugosity should be integrated via pcA. Scarid density and conspific density should also be combined, or one should be omitted.

```{r,echo=FALSE}
pairs(~cca_cover+lc_cover+ma_cover+ma_canopy+ta_cover+ta_canopy+rugosity, data = sum_id,
      lower.panel = panel.smooth, upper.panel = panel.cor, 
      pch=20, main ="Benthic Community Variables", na.action = na.omit)
```

All benthic variables are highly correlated. I can omit cca_cover and ma_cover (accounted for by ma_canopy). The rest should be combined via pcA. First I will assess remaining correlations between selected benthic and fish community variables:

```{r,echo=FALSE}
pairs(~scar_bm+scar_den+carn_bm+lc_cover+ta_cover+ma_canopy+ta_canopy+rugosity, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish and Benthic Variables",na.action = na.omit)
```

Unsure how to proceed here if I do end up using both benthic and fish community variables as predictors, because some are corellated (esp. scarid_bm.s and all benthic variables....). It looks like I would integrate scar_bm and carn_bm into a pcA with all other benthic/rugosity variables. Variables scar_den and consp_den are relatively independent of all other variables (except each other).

Current outline for site-level pcA tasks:
If benthic=predictor:
- lc_cover + ta_cover + ma_canopy + ta_canopy + rugosity + scar_bm + carn_bm
- scar_den + consp_den
If benthic=response:
- rugosity + scar_bm + carn_bm
- lc_cover + ta_cover + ma_canopy + ta_canopy
- scar_den + consp_den

## Fish-level response variables: grazing behaviors

### Distributions
```{r,echo=FALSE,warning=FALSE}
grazing <- sum_id %>% 
  select(g_frac,fr,br,for_dur,for_bites) %>% 
  gather() %>% 
  na.omit()

ggplot(grazing,aes(value)) + 
  facet_wrap(~ key, scales = "free") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

### Corellations
```{r,echo=FALSE}
cor=cor(sum_id %>% 
          select(g_frac,fr,br,for_dur,for_bites), use="pairwise.complete.obs")
corplot <- corrplot(cor,title = "Correlation Plot", method = "square", outline = T, addgrid.col = "darkgray", order="hclust", mar = c(4,0,4,0), addrect = 4, rect.col = "black", rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", tl.cex = .5, cl.cex = .5)
```

```{r,echo=FALSE}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

Variable for_bites is correlated with nearly all other variables, so will be omitted. Variables g.farc, fr. and for_dur are still correlated and should be assessed with a pcA. Variable br can be a standalone variable.

Grazing pcA:
- g_frac + fr + for_dur
- (br as standalone)

# Exploratory bivariate plots

## Grazing as a function of fish length
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = br, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r, echo=FALSE,warning=FALSE, eval=FALSE}
ggplot(sum_id, aes(x = length_cm, y = g_frac, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  ggtitle("Feeding rate by fish length") + 
  labs(y = "Time Fraction", x = "Length (cm)") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```

## Grazing as a function of scarid biomass

```{r}
ggplot(sum_id %>% filter(island!="Barbuda"), aes(x = scar_bm,y=fr)) +
  facet_grid(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = island)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = fr, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Feeding rate by phase and scarid biomass") +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = br, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bite rates (while grazing) by scarid biomass") +
  labs(y = "Bite rate (bites/sec)", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = g_frac, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Fraction of time spent grazing by phase and scarid biomass") +
  labs(y = "Time Fraction", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id, aes(x = scar_bm, y = for_bites, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  ggtitle("Bites per foray by phase and scarid biomass") +
  labs(y = "Bites", x = "Scarid Biomass (g/100m2)") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```

## Competitive interactions

```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean, shape=island)) +
  geom_point(size=3) +
  ggtitle("Interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean-int_scar.se, ymax=int_scar.mean+int_scar.se))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_stop, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. viride interaction frequency as a function of scarid biomass") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Scarid Biomass (g/100m2)", y = "Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar.mean_stop-int_scar.se_stop, ymax=int_scar.mean_stop+int_scar.se_stop))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar.mean_qup, shape=island)) +
  geom_point(size=3) +
  ggtitle("S. vetula interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") + 
  geom_errorbar(aes(ymin=int_scar.mean_qup-int_scar.se_qup, ymax=int_scar.mean_qup+int_scar.se_qup))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(sum_id,aes(x=scar_bm,y=int_scar)) +
  facet_grid(.~species)+geom_point(shape=19) +
  geom_smooth(method=lm) +
  ggtitle("Competitive interactions as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

# PCAs -->
*Do I need to normalize my variables pre-pcA? Currently just scaled, except for grazing response variables which are log-transformed* -->

## Predictors

```{r}
pca_pred <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              select(lc_cover,ta_cover,ma_canopy,ta_canopy,rugosity,scar_bm,scar_den,tot_bm)),
  center = T,scale = T)

summary(pca_pred)
plot(pca_pred,type="l")

autoplot(pca_pred, data = sum_site, colour = 'island',
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

# add PC1 and PC2 as variables in dataset
sum_site_pca1 <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  clean_names(case="old_janitor")
sum_id_pca1 <- sum_id %>% left_join(select(sum_site_pca1,site,pc1,pc2), by = "site")
write.csv(sum_id_pca1,file=here("data/sum_id_pca1.csv"))
```

## Response variables

```{r}
pca_resp <- prcomp(na.omit(sum_id %>% 
              select(g_frac,fr,br)), 
  center = T,scale = T)

summary(pca_resp)
plot(pca_resp,type="l")

autoplot(pca_resp, data = sum_id, colour = "island",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black', 
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("navy", "darkseagreen", "slategray2")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

# autoplot doesn't like when I add it for_dur or for_bites (NA values?)
```
# Model trials

-site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)
-fish-level predictors: species, phase, length
--eventually run separately for different species
--species*scar_bm interaction?
-random effects: island
-response variables: g_frac, br (?), and fr (run separately)

## Basic lm 
```{r}
lm_fr <- lm(fr~species+phase+length_cm+pc1+pc2, data=filter(sum_id_pca1, species_code!="rbp"))
summary(lm_fr) 
```

## Mixed effects models
- random effect: island

```{r}
lme_fr <- lme(fr~phase+length_cm+species+pc1+pc2,data=filter(sum_id_pca1, species_code!="rbp"),random=~1|island, na.action = na.omit)
summary(lme_fr)
```

## GAMs
```{r, echo=FALSE}
gam<-gam(fr~species+phase+s(length_cm)+s(scar_bm)+s(pc1)+s(pc2), family=gaussian(link=identity),sum_id_pca1=data, method="REML")
summary(gam)
plot(gam,pages=1)
```

## GAMMs

```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="qup"), random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,message=FALSE,warning=FALSE}
gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2), family=gaussian(link=identity),data=filter(sum_id_pca1, species_code=="stop"), random=list(island=~1),method="REML")
summary(gamm$gam)
plot(gamm$gam,pages=1)
```

```
------------------------
To Do as of Nov. 7
* ~~incorporate pcs~~
* boosted regression trees ecosphere 2017 adrians paper
* ~~simpler graphs with just herb density x axis - grazing behaviors y, facet by species~~
* ~~species*density as fixed effect~~
* species as random effect
* ~~break out species and run separately~~