### 2018-11-15 r/Analysis.Rmd
11/14

---
title: "Figures + Analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Outline of selected figures and analyses:

- Grazing differences between islands: (site or island level?)
-- ANOVAs to test statistical significance of differences here
- Total grazing impact of site-level biomass: measured vs. reported grazing rates (add line or color by feeding rate)
- Something looking at relationship with algal index vs. scarid biomass  
- SEM?

Notes: 
- re: lms/gamms: don't focus on these for now, the data has so many colinearities... if anything, use full PCA in model analysis
- Exclude Barbuda from this analysis...?

```{r}
# set up
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(startR)
library(here)
library(tidyverse)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
library(ggfortify)
library(ggsignif) # boxplot stats
library(EnvStats) # boxplot stats
library(car)
library(nlme)
library(gamm4)
```
```{r}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(length_cm>=15)
fish <- read.csv(here("data","fish.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Island - level differences in grazing rates

```{r}
# Violin plot of inter-island differences
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), fr)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5))

# Length-weight relationships between islands
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), 
       aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))

# Calculating length-weight coefficients

fr_anu_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "stop"))
summary(fr_anu_vir) # fr(Antigua, S. viride, initial) = 445-8*L

fr_anu_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "qup"))
summary(fr_anu_vet) # fr(Antigua, S. vetula, initial) = 1030-12*L

fr_bon_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "stop"))
summary(fr_bon_vir) # fr(Bonaire, S. viride, initial) = 1057-17*L

fr_bon_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "qup"))
summary(fr_bon_vet) # fr(Bonaire, S. vetula, initial) = 1347+10*L
```

fr(Antigua, *S. viride*, initial) = 445-8*L
fr(Antigua, *S. vetula*, initial) = 1030-12*L
fr(Bonaire, *S. viride*, initial) = 1057-17*L
fr(Bonaire, *S. vetula*, initial) = 1347+10*L

## Comparison to published equations

## Predicted feeding rates and bite sizes from standard equations
Bite sizes as predicted by Bruggemann/Bozec:
bs(*S. vetula*) = 4.013*10^-4*L^2
bs(*S. viride*) = 5.257*10^-4*L^2

Feeding rates as published by Bruggemann/Bozec
fr(Published, *S. viride*, initial) = 1089-17*L-56 
  *(extra term because phase modifier is not used here (i=1))*
fr(Published, *S. vetula*, initial) = 3329-33*L

## adding predicted feeding rates from observed relationships

```{r}
fish_fr <- fish %>% 
  filter(species_code %in% c("stop", "qup") & island %in% c("Antigua","Bonaire")) %>%
  mutate(
    fr_pers = ifelse(island == "Antigua" & species_code == "stop", 445-8*length,
                     ifelse(island == "Antigua" & species_code == "qup", 1030-12*length,
                            ifelse(island == "Bonaire" & species_code == "stop", 
                                   1057-17*length, 1347+10*length))),
    fr_pub = ifelse(species_code == "stop", 1089-17*length-56, 3329-33*length),
    bs = ifelse(species_code == "stop", 5.257*10^-4*length^2, 4.013*10^-4*length^2),
    ag_pers = fr_pers*bs,
    ag_pub = fr_pub*bs
)

ag_site <- fish_fr %>% group_by(site,island) %>%
  summarize(personal=sum(ag_pers)/1000,published=sum(ag_pub)/1000)
ag_melt <- ag_site %>% as.data.frame() %>% 
  gather(source, area_grazed, -c(site,island))

ggplot(filter(ag_melt, island == "Antigua"),
      aes(x=reorder(site, -area_grazed), y=area_grazed)) +
        geom_bar(aes(fill = source), position = "dodge", stat = "identity") +
        scale_fill_manual(values = c("black", "slategray2")) +
        labs(title=expression(Predicted~grazing~impact~on~Antiguan~reefs~(m^2/hr))) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```