### 2018-11-15 r/Sites.Rmd

---
title: "Fish & Benthic Community data: Antigua, Barbuda, & Bonaire - Aug. 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Wrangling of fish, benthic, and rugosity data from Antigua, Barbuda, and Bonaire to put in usable site-level format.

Final outputs:
* "site.csv": site-level summary data of fish, benthic, rugosity, and human impact variables
* "sp_site.csv": site-level fish data by species (ultimately used to merge conspecific data with fish follows df)

# Fish community data

```{r, echo=F,message=F,warning=F}
# set up
library(startR)
library(here)
library(janitor)
library(tidyverse)
```
```{r, echo=F,message=F,warning=F}
# import files
ab_fish <-read.csv(here("data","ab_fish.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  uncount(number) %>% # expand to replicate rows if multiple fish were recorded
  select(-c(time,depth_ft,surveyor,date)) 
bon_herb <- read.csv(here("data","bon_herb.csv"), stringsAsFactors=F) %>% # this file is already expanded
  janitor::clean_names(case="old_janitor") %>%
  filter(date==2017) %>% 
  select(-date) %>%
  rename(island = region)
bon_carn <- read.csv(here("data","bon_carn.csv"), stringsAsFactors=F) %>% 
  janitor::clean_names(case="old_janitor") %>%
  rename(island = region) %>%
  uncount(number) %>%
  select(-c(diver,date))
fish <- bind_rows(bon_herb,bon_carn) %>% 
  filter(site %in% c("Oil Slick","Karpata","Front Porch","18th Palm")) %>%
  bind_rows(ab_fish) %>%
  mutate(
    biomass = as.numeric(biomass),
    species_code = tolower(species_code)) %>%
  replace(. == "", NA)

ab_benthic <- read.csv(here("data","ab_benthic.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor") %>%
  mutate(height=as.numeric(height))
ab_rugosity <- read.csv(here("data","ab_rugosity.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")
bon_benthic <- read.csv(here("data","bon_benthic.csv"),stringsAsFactors = F) %>%
  janitor::clean_names(case="old_janitor")
human <- read.csv(here("data","site_human.csv"), stringsAsFactors=F) %>%
  janitor::clean_names(case="old_janitor")
```

## Total fish population, by site
```{r,message=F,warning=F}
fish_tran <- fish %>% 
  filter(biomass != 0) %>% 
  group_by(island,site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2, 
    den_t=n()/1.2, 
    l_t=mean(length)
    )

fish_site <-fish_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(), 
    tot_bm=mean(bm_t), 
    tot_bm_se=(sd(bm_t))/sqrt(n), 
    tot_den=mean(den_t), 
    tot_den_se=(sd(den_t))/sqrt(n), 
    tot_l=mean(l_t), 
    tot_l_se=(sd(l_t))/sqrt(n)
    ) %>% 
  select(-n) # average across transects within each site

boxplot(bm_t~site, las=2, data = fish_tran)
```

## Total fish populations, by species and site
```{r,message=F,warning=F}
# Because not all species were present at each site/transect, I need to create template to add 0 values for biomass and density in transects where they are absent.
exp_tran <- fish %>% 
  group_by(site,transect) %>% 
  summarize() # all possible site/transect combinations
exp_sp <- fish %>% tidyr::expand(nesting(species,family),site) # all possible species/site combinations
template_sp <- full_join(exp_sp,exp_tran,by="site") # species/site/transect combinations

# Biomass and density, by site and species:
sp_tran <- fish %>% 
  filter(biomass != 0) %>% 
  group_by(site,transect,species) %>% 
  summarize(
    bm_t=sum(biomass)/1.2, 
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(template_sp, by=c("site","transect","species")) %>% # join transect-level data with template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t) # missing biomass and density values should be 0

sp_site<-sp_tran %>% 
  group_by(site,species,family) %>% 
  summarize(
    n=n(), # sample size for biomass and density includes 0 values
    l_n=sum(!is.na(l_t)), # only count non-0 values in sample size for length
    bm=mean(bm_t), 
    bm_se=(sd(bm_t))/sqrt(n), 
    den=mean(den_t), 
    den_se=(sd(den_t))/sqrt(n),
    l=mean(l_t, na.rm = T),
    l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>%
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

```

## Total fish populations, by functional group and site
```{r,message=F,warning=F}
# Herbivores
herb_tran<- fish %>% 
  filter(family %in% c("Scaridae","Acanthuridae"),
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran, by=c("site","transect")) %>% # join with transect-level template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t) # missing biomass and density values should be 0

herb_site<-herb_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    herb_bm=mean(bm_t), 
    herb_bm_se=(sd(bm_t))/sqrt(n), 
    herb_den=mean(den_t), 
    herb_den_se=(sd(den_t))/sqrt(n), 
    herb_l=mean(l_t, na.rm = T),
    herb_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Acanthurids

acan_tran <- fish %>% 
  filter(family=="Acanthuridae",
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>%
  right_join(exp_tran, by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

acan_site<-acan_tran %>% group_by(site) %>% 
  summarize(
    n=n(), 
    l_n=sum(!is.na(l_t)),
    acan_bm=mean(bm_t),
    acan_bm_se=(sd(bm_t))/sqrt(n),
    acan_den=mean(den_t),
    acan_den_se=(sd(den_t))/sqrt(n),
    acan_l=mean(l_t, na.rm = T),
    acan_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Carnivores
carn_tran <- fish %>% 
  filter(family %in% c("Serranidae","lutjanidae","Carangidae"),
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran,by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

carn_site<-carn_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    carn_bm=mean(bm_t),
    carn_bm_se=(sd(bm_t))/sqrt(n),
    carn_den=mean(den_t),
    carn_den_se=(sd(den_t))/sqrt(n),
    carn_l=mean(l_t, na.rm = T),
    carn_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

```

## Scarid populations, by site

```{r,message=F,warning=F}
# Total scarids
scar_tran <- fish %>% 
  filter(family=="Scaridae",
         biomass != 0) %>% 
  group_by(site,transect) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(exp_tran,by=c("site","transect")) %>%
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

scar_site <- scar_tran %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    scar_bm=mean(bm_t),
    scar_bm_se=(sd(bm_t))/sqrt(n),
    scar_den=mean(den_t),
    scar_den_se=(sd(den_t))/sqrt(n),
    scar_l=mean(l_t, na.rm = T),
    scar_l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>% 
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

# Scarids by species (join wide S. vetula, S. viride, and S. aurofrenatum density, biomass, and length data)

scar_site_sp <- sp_site %>% 
    ungroup() %>%
    filter(species=="Sparisoma viride") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("vir_",.)) %>%
  left_join(sp_site %>% 
    ungroup() %>%
    filter(species=="Scarus vetula") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("vet_",.)),
    by="site"
  ) %>%
  left_join(sp_site %>% 
    ungroup() %>%
    filter(species=="Sparisoma aurofrenatum") %>% 
    select(-c(species,family)) %>%
    rename_at(vars(2:7), ~ paste0("aur_",.)),
    by="site"
  )

# Scarids by phase (should this be by species, too?)
exp_ph <- fish %>% tidyr::expand(phase,site)
template_ph <- full_join(exp_ph,exp_tran,by="site") # all site/transect/phase combinations

ph_tran <- fish %>% 
  filter(family=="Scaridae",
         biomass != 0) %>% 
  group_by(site,transect,phase) %>% 
  summarize(
    bm_t=sum(biomass)/1.2,
    den_t=n()/1.2,
    l_t=mean(length)
    ) %>% 
  right_join(template_ph, by=c("site","transect","phase")) %>% # join transect-level data with template
  mutate_each(funs(replace(.,is.na(.),0)), -l_t)

ph_site <- ph_tran %>% 
  group_by(site,phase) %>% 
  summarize(
    n=n(),
    l_n=sum(!is.na(l_t)),
    bm=mean(bm_t),
    bm_se=(sd(bm_t))/sqrt(n),
    den=mean(den_t),
    den_se=(sd(den_t))/sqrt(l_n),
    l=mean(l_t, na.rm = T),
    l_se=(sd(l_t, na.rm = T))/sqrt(l_n)
    ) %>%
  select(-c(n,l_n)) %>%
  replace(. == "NaN",NA)

scar_site_ph <- ph_site %>% 
    ungroup() %>%
    filter(phase=="j") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("juv_",.)) %>%
  left_join(ph_site %>% 
    ungroup() %>%
    filter(phase=="i") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("init_",.)),
    by="site"
  ) %>%
  left_join(ph_site %>% 
    ungroup() %>%
    filter(phase=="t") %>% 
    select(-c(phase)) %>%
    rename_at(vars(2:7), ~ paste0("term_",.)),
    by="site"
  )
```

# Benthic & rugosity data

## Percent cover and canopy heights

```{r, echo=F,message=F,warning=F}
# Because not all categories were present within each meter, I need to create template to add 0 values in meters where they were ab_fishsent.
exp_meter <- ab_benthic %>% 
  group_by(site,transect,meter) %>% 
  summarize()
exp_cat <- ab_benthic %>% tidyr::expand(nesting(code),site)
template_bent <- full_join(exp_meter,exp_cat,by="site")

# Then I can join benthic data with this template, and replace empty rows with 0 values
bent_meter <- ab_benthic %>% 
  group_by(site,transect,meter,code) %>% 
  summarize(count_m=n()) %>% 
  right_join(template_bent, by=c("site","transect","meter","code")) %>% # join transect-level data with template
  replace(is.na(.),0) # convert NAs to 0 biomass/density in transects where a species is not present

# Calculating percent cover on potential substrate (i.e. not sand), and remove meters that were 100% sand cover (no potential susbtrate)

sand_meter <- bent_meter %>% 
  filter(code=="sand") %>% 
  rename(count_sand=count_m) %>% 
  select(-code) 
bent_meter <- left_join(bent_meter,sand_meter,by=c("site","transect","meter")) %>%
  mutate(cover_m=100*count_m/(10-count_sand)) %>% 
  filter(!is.nan(cover_m)) %>% filter(code!="sand") 
bent_tran <- bent_meter %>% 
  group_by(site,transect,code) %>% 
  summarize(cover_t=mean(cover_m))
bent_site <- bent_tran %>% 
  group_by(site,code) %>% 
  summarize(
    n=n(),
    cover=mean(cover_t),
    cover_se=sd(cover_t)/sqrt(n)
    ) %>% 
  select(-n)

# Canopy heights

can_site <- ab_benthic %>% 
  filter(height>=1) %>% 
  group_by(site,transect,meter,code) %>% 
  summarize(canopy_m=mean(height)) %>% 
  group_by(site,transect,code) %>% 
  summarize(canopy_t=mean(canopy_m)) %>% 
  group_by(site,code) %>% 
  summarize(
    n=n(),
    canopy=mean(canopy_t),
    canopy_se=sd(canopy_t)/sqrt(n)
    ) %>% 
  select(-n)

# Spread into wide site-level format, join with percent cover data in benthic.wide

spread_n <- function(df, key, value) {
    keyq <- rlang::enquo(key)
    valueq <- rlang::enquo(value)
    s <- rlang::quos(!!valueq)
    df %>% gather(variable, value, !!!s) %>%
        unite(temp, !!keyq, variable) %>%
        spread(temp, value)
}

bent_site <- bent_site %>% 
      spread_n(code, c(cover, cover_se)) %>%
  left_join(can_site %>% 
            spread_n(code, c(canopy,canopy_se))
  ) %>% 
  janitor::clean_names(case="old_janitor") %>%
  select(cca_cover, cca_cover_se, lc_cover, lc_cover_se, ma_cover, ma_cover_se, ma_canopy, ma_canopy_se, ta_cover, ta_cover_se, ta_canopy, ta_canopy_se)

```

## Rugosity
```{r, echo=F,message=F,warning=F}
rug_site <- ab_rugosity %>% 
  group_by(site,transect) %>% 
  summarize(rugosity_t=mean(rugosity_cm)/100) %>% 
  group_by(site) %>% 
  summarize(
    n=n(),
    rugosity=mean(rugosity_t),
    rugosity_se=sd(rugosity_t)/sqrt(n)
    ) %>% 
  select(-n) %>% 
  rbind(c("Pallaster West",1.7006,NA)) %>% # missing data for Pallaster West, will use mean of Pallaster East and Mid because reef structure across all three was extremely similar
  mutate(
    rugosity=as.numeric(rugosity),
    rugosity_se=as.numeric(rugosity_se)
  )

bent_site <- bent_site %>% 
  left_join(rug_site,by="site") %>% 
  bind_rows(bon_benthic) %>%
  mutate(ma_ind = ma_canopy*ma_cover,
         ta_ind = ta_canopy*ta_cover)
```

# Integrating and exporting .csvs
```{r, echo=F,message=F,warning=F}
site <- fish_site %>% 
  left_join(herb_site,by="site") %>% 
  left_join(acan_site,by="site") %>% 
  left_join(scar_site,by="site") %>% 
  left_join(scar_site_sp,by="site") %>% 
  left_join(scar_site_ph,by="site") %>% 
  left_join(carn_site,by="site") %>%
  left_join(bent_site,by="site") %>%
  left_join(human,by="site")

write.csv(site,file=here("data/site.csv"))
write.csv(sp_site,file=here("data/sp_site.csv"))
write.csv(fish,file=here("data/fish.csv")) # for predicting grazing impact of observed fish communities at each site (see Analysis.rmd)
```