### 2018-11-24 r/Models.Rmd

---
title: "Scarid behavior GAM/GAMMS"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(startR)
library(here)
library(tidyverse)
library(mgcv)
library(car)

sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

# Examining predictor variables
Notes: in models using benthic-only PCA, scarid biomass and scarid density are highly collinear with benthic PCA variables (understandably). I don't think I can justify separating them, even though they are my ultimate variables of interest. I can use PCA loadings to interpret model outcomes. Carnivore biomass and spearfishing presence is also highly correlated with scarid biomass and thus benthic PCA variables - I am excluding it for now.

```{r}
pca_all <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
  center = T,scale = T)
autoplot(pca_all, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "salmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

pca_bent <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity)),
  center = T,scale = T)

sum_site <- sum_site %>% 
  # filter(island!="Barbuda") %>%
  bind_cols(as.data.frame(pca_bent$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_bent=PC1,
         pc2_bent=PC2) %>% 
  bind_cols(as.data.frame(pca_all$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")

```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm + pc1_bent + pc2_bent, data=sum_site,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish & Benthic Community Variables", na.action = na.omit)
```

```{r}
m_data <- sum_id_pca %>% filter(species_code!="rbp")
vif(lm(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data))
vif(lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
        random=~1|island, 
        na.action = na.omit
        ))
```

Confirming I can't use scarid biomass and/or density as separate from benthic PCA variables. VIF of scar_bm is just slightly >3. Removing rugosity from benthic PCA reduces scar_bm ~ pc1 correlation slightly, but still >0.7.

BUT: VIF done on LME where random=site or random=island shows values <3 when using benthic PCA values + scarid biomass... *stay tuned...*

# Model as LME first
```{r}
lme1 <- lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme1)
vif(lme1)
```

# Switching to gamm, checking model validity

Notes:  
- GLS: option for model when homogeneity of variances is not satisfied, can add weights for residuals. Keep reading Zuur et al. and decide if I need to go beyond this (e.g. GAMM)

```{r}
m1 <- gam(fr ~ factor(species) + s(length_cm, bs = "cs") + s(pc1_all, bs = "cs") + s(pc2_all, bs = "cs"), data=m_data)
summary(m1)
plot(m1)
gam.check(m1)

gls1 <- gam(fr ~ species + length_cm + pc1_all + pc2_all, data=m_data)
summary(gls1)

e <- resid(m1)
coplot(e ~ length_cm | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc1_all | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc2_all | species, data = m_data) # inconsistent res.

m2 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) +s (pc2_all) +
          s(length_cm, by = as.numeric(species_code=="qup")), 
          data = m_data)
summary(m2)
gam.check(m2)

m3 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc1_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m3)
gam.check(m3)

m4 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc2_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m4)
gam.check(m4)

m5 <- gam(fr ~ factor(species_code) + s(length_cm) + s(pc1_all) + s(pc2_all) +
          s(pc1_all, by = as.numeric(species_code == "qup"))+
          s(pc2_all, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m5)
plot(m5)
gam.check(m5)

```

Notes:  
- Violates assumption of homogeneity of variance (increasing variance as predictor values increase)
- Satisfies assumption of normality

```{r}
gamm<-gamm(fr~phase+s(length_cm)+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_id_pca1, species_code=="qup"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r}
gamm1<-gamm(fr_mean~phase+species+s(pc1)+s(pc2)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_ssp_pca1, species_code!="rbp" & island!="Barbuda"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm1$gam)
gamtabs(gamm1)
AIC(gamm1$lme)
plot(gamm1$gam,pages=1)

```