### 2018-11-26 r/Data_prodding.Rmd

---
title: "Investigation of scarid data"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true

---

Notes 11/11:
* figure out dcast issues
* replacing NA/0s within pipe

```{r,message=FALSE,warning=FALSE, echo=FALSE}
# set up
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(startR)
library(here)
library(tidyverse)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
library(ggfortify)
library(ggsignif) # boxplot stats
library(EnvStats) # boxplot stats
library(car)
library(nlme)
library(gamm4)
library(xtable)
```
```{r,message=FALSE,warning=FALSE, echo=FALSE}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  filter(length_cm>=15)
sum_site_sp_ph <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X) 
fish <- read.csv(here("data","fish.csv"), stringsAsFactors=F) %>%
  select(-X) 
```

# Data overview:
- Fish follows (2 min) conducted at sites in Antigua (6), Barbuda (3) and Bonaire (4) from March-August 2017
- Follows tracked time spent grazing, bite rates, and competitive interactions (among scarids and with damselfish)
- Follows targeted *Sparisoma viride*, *Scarus vetula*, and *Sparisoma aurofrenatum* of both initial and terminal phases 
  + *Sparisoma aurofrenatum* were not followed in Bonaire, but added in Antigua and Bonaire due to low abundances of other two species. Some sites in Antigua and Barbuda had no/low represenation from terminal phase *viride* and *vetula*
  + Most eventual analyses will likely focus exclusively on initial phase (standardized size window) *S. viride* and *S. vetula*
- site level factors (benthic, fish, and rugosity) assessed at each of the 13 sites

```{r}
sample <- group_by(sum_id, site, species, species_code, phase, island) %>% summarize(n=n())

ggplot(sample, aes(x=factor(site, levels = sum_site$site), y=n)) +
        facet_grid(phase~species) +
        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
        scale_fill_manual(values = c("gray", "slategray2", "black")) +
        labs(title="Total sample") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}
ggplot(filter(sample, species_code!="rbp" & phase == "i"),
      aes(x=factor(site, levels = sum_site$site), y=n)) +
        facet_grid(.~species) +
        geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
        scale_fill_manual(values = c("gray", "slategray2", "black")) +
        labs(title="Sample subset") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Need to determine appropriate size range for comparison. Because they are not evenly distributed (i.e. larger fish in Bonaire, smaller in Barbuda), I will likely want to compare length-feeding relationships as opposed to pooled averages

```{r}
ggplot(filter(sum_id, species_code!="rbp" & phase == "i" & length_cm<=30 & length_cm>=15), aes(x = length_cm)) +
  facet_grid(.~species) +
  geom_density(aes(fill = island), alpha=0.4) +
  labs(title="Size distributions within sample subset") +
  theme_bw() 
```

# Examing variable distributions and relationships

## Predictor variables
Potential predictor variables are site-level fish, benthic, and rugosity values. These are likely correlated to one another, and I need to determine which ones I ultimately want to use (if modeling behavioral responses via any multivariate regressions). I can also move to SEM if I want to keep multiple correlated predictors.

First, check distribution of predictor variables of interest: not very normally distributed...
```{r,warning=FALSE,echo=FALSE}
pred <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn30_bm,carn30_den,
         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity,spear_daily) %>% 
  gather() %>% 
  na.omit()

ggplot(pred,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

```{r,echo=FALSE}
pred <- sum_id %>% select(tot_bm,tot_den,scar_bm,scar_den,consp_bm,consp_den,
                           carn30_bm,carn30_den,cca_cover,lc_cover,
                           ma_cover, ma_canopy,ta_cover,ta_canopy,rugosity)
c = cor(pred, use="pairwise.complete.obs")
corplot <- corrplot(c, type = "upper", order = "hclust")
```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~consp_l+scar_bm+rugosity+ma_ind+ta_canopy+ta_cover, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish & Benthic Community Variables", na.action = na.omit)
```

Variable selection notes:
- excluding both carnivore variables as they are highly correlated with scarid biomass and total biomass, eventually I could make these more nuanced by distinguishing actual predators, but right now I don't think it reflects actual predator populations of >15cm parrotfish
- rugosity is highly correlated with turf cover, and scarid density
- scarid density: removing for now, because I think it was a bit skewed from Barbuda juveniles
- could eventually use consp. scarid length as another indicator of overfishing?

PCA to visualize variable relationships:

```{r}
pca_pred <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity, scar_den, 
                     scar_bm, spear_daily)),
  center = T,scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()
```

PCA for correlated (benthic only?) variables:

```{r}
pca_bent <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, rugosity)), 
  center = T,scale = T)
summary(pca_bent)
plot(pca_bent,type="l")

autoplot(pca_bent, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

ggsave(filename = here("figs/pca_bent.png"))

# add PC1 and PC2 as variables in dataset
sum_site <- sum_site %>% 
  # filter(island!="Barbuda") %>%
  bind_cols(as.data.frame(pca_bent$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_bent=PC1,
         pc2_bent=PC2) %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
write.csv(sum_id_pca, file=here("data/sum_id_pca.csv"))

sum_ssp_pca <- sum_site_sp_ph %>%
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
write.csv(sum_ssp_pca, file=here("data/sum_ssp_pca.csv"))
```

## Response variables
Fish-level grazing behaviors (as well as competitive interaction frequency)

### Distributions
```{r,echo=FALSE,warning=FALSE}
grazing <- sum_id %>% 
  select(g_frac,fr,br,for_dur,for_bites) %>% 
  gather() %>% 
  na.omit()

ggplot(grazing,aes(value)) + 
  facet_wrap(~ key, scales = "free") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

```{r,echo=FALSE}
pairs(~g_frac+fr+br+for_dur+for_bites, data=sum_id,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Grazing Response Variables",na.action = na.omit)
```

Variable selection notes:
- for_bites is correlated with fr and for_dur, but I will play around with keeping it for now.

# Grazing boxplots

```{r, message=FALSE,warning=FALSE,echo=FALSE}
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(x=island, y=fr)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom="point", shape=5, size=3) + 
  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  theme(strip.text = element_text(face = "italic")) + 
  stat_n_text()
ggsave(filename = here("figs/bp_fr.png"))

ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), fr)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(filename = here("figs/vp_fr.png"))

ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(site), fr)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90, hjust = 1)) 
```

```{r}
vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
aov_fr_vet <- aov(fr ~ island, data = vet)
summary.aov(aov_fr_vet) # significant differences
TukeyHSD(aov_fr_vet)
plot(aov_fr_vet, 2) # normally distributed
leveneTest(fr ~ island, data = vet) # homogeneity of variance
```
```{r}
vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
aov_fr_vir <- aov(fr ~ island, data = vir, white.adjust=T) 
summary.aov(aov_fr_vir) # significant differences
TukeyHSD(aov_fr_vir)
plot(aov_fr_vir, 2) # normally distributed
leveneTest(fr ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

```{r}
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), g_frac)) + 
  geom_boxplot() + 
  labs(y = "Fraction of time", x = "",title="Grazing time budgets (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(here("figs/bp_gfrac.png"))

ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), g_frac)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Fraction of time", x = "",title="Grazing time budgets (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5))
ggsave(here("figs/vp_gfrac.png"))
```

```{r}
vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
aov_g_vet <- aov(g_frac ~ island, data = vet, white.adjust = T)
summary.aov(aov_g_vet) # significant differences
TukeyHSD(aov_g_vet)
plot(aov_g_vet, 2) # normally distributed
leveneTest(g_frac ~ island, data = vet) # heteroskedastic - add white.adjust = 2
```
```{r}
vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
aov_g_vir <- aov(g_frac ~ island, data = vir, white.adjust=T) 
summary.aov(aov_g_vir) # significant differences
TukeyHSD(aov_g_vir)
plot(aov_g_vir, 2) # normally distributed
leveneTest(g_frac ~ island, data = vir) # homogeneity of variance
```

```{r}
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), br)) + 
  geom_boxplot() + 
  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(here("figs/bp_br.png"))

ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), br)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Bite Rate (bites/sec)", x = "", title = "Bite rate while grazing (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(here("figs/vp_br.png"))
```

```{r}
vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
aov_br_vet <- aov(br ~ island, data = vet)
summary.aov(aov_br_vet) # significant differences
TukeyHSD(aov_br_vet)
plot(aov_br_vet, 2) # normally distributed
leveneTest(br ~ island, data = vet) # homogeneity of variance
```
```{r}
vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
aov_br_vir <- aov(br ~ island, data = vir, white.adjust=T) 
summary.aov(aov_br_vir) # significant differences
TukeyHSD(aov_br_vir)
plot(aov_br_vir, 2) # normally distributed
leveneTest(br ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

```{r}
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), for_bites)) + 
  geom_boxplot() + 
  labs(y = "Bites per foray", x = "", title = "Bites per foray (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(here("figs/bp_fordur.png"))

ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), for_bites)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Bites per foray", x = "", title = "Bites per foray (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(here("figs/vp_fordur.png"))

vet <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="qup")
aov_for_bites_vet <- aov(for_bites ~ island, data = vet)
summary.aov(aov_for_bites_vet) # significant differences
TukeyHSD(aov_for_bites_vet)
plot(aov_for_bites_vet, 2) # normally distributed
leveneTest(for_bites ~ island, data = vet) # homogeneity of variance

vir <- sum_id %>% filter(phase == "i" & length_cm >= 16 & length_cm <= 24 & species_code=="stop")
aov_for_bites_vir <- aov(for_bites ~ island, data = vir, white.adjust=T) 
summary.aov(aov_for_bites_vir) # significant differences
TukeyHSD(aov_for_bites_vir)
plot(aov_for_bites_vir, 2) # normally distributed
leveneTest(for_bites ~ island, data = vir) # heteroskedastic - add white.adjust=TRUE
```

# Exploratory bivariate plots

## Grazing as a function of fish length

```{r}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), 
       aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
  theme_bw() + 
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
ggsave(here("figs/sp_fr_length.png"))
```

```{r}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & site!="Turtle Bay"), aes(x = length_cm, y = fr, colour = site)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = br, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Bite Rate (bites/sec)", x = "Length (cm)", title = "Bite rate while grazing") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```

*Note: remove G1 grazing instances here?*

```{r}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = length_cm, y = g_frac, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Time Fraction", x = "Length (cm)", title = "Fraction of time spent grazing") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```

## Grazing as a function of site traits

```{r}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm,y=fr)) +
  facet_grid(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  labs(y="Feeding rate (bites/hr)", x="Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r,echo=FALSE,warning=FALSE}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = scar_bm, y = fr, colour = island)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))
```
```{r,echo=FALSE,warning=FALSE}
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), aes(x = scar_bm, y = fr, colour = phase)) +
  facet_grid(. ~ species) +
  geom_point(shape = 18) +
  geom_smooth(method = lm, fullrange=F) +
  labs(y = "Feeding rate (bites/hr)", x = "Scarid Biomass (g/100m2)", title = "Hourly feeding rate") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("gray40", "aquamarine3"))
```

Notes: 
- scarid biomass is not the best predictor once I account for differences between my samples in terms of the sizes of fish I was sampling. I think the grazing/length relationships are much stronger.  
- restraining sample size to length windows lowers sample size and makes trends much less pronounced - esp. for phase differences  
- reducing sample to individual phase only also blurs trends

```{r}
ggplot(filter(sum_id_pca, species_code != "rbp" & island != "Barbuda" & length_cm>=16 & length_cm<=24), aes(x = ma_ind,y=fr)) +
  facet_grid(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  labs(y="Feeding rate (bites/hr)", x="Macroalgal index", title = "Hourly feeding rate") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

## Competitive interactions

```{r,echo=FALSE,warning=FALSE}
ggplot(sum_site,aes(x=scar_bm,y=int_scar_mean, shape=island)) +
  geom_point(size=3) +
  ggtitle("Interaction frequency as a function of scarid biomass") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x="Scarid Biomass (g/100m2)", y="Interactions with other scarids (per 2 min follow)") +
  geom_errorbar(aes(ymin=int_scar_mean-int_scar_se, ymax=int_scar_mean+int_scar_se))
```

# Model trials

- site-level predictors: scar_bm,scar_den,carn_bm,benthic (pc1,pc2)  
- fish-level predictors: species, phase, length  
  + eventually run separately for different species  
  + species*scar_bm interaction?  
- random effects: island  
- response variables: g_frac, br (?), and fr (run separately)

## Mixed effects models
- random effect: island or site?

```{r}
lme_fr <- lme(fr~phase+length_cm+species+scar_bm+pc1_bent+pc2_bent,
              data=filter(sum_id_pca, species_code!="rbp"),
              random=~1|island, # do I need random slope too?
              na.action = na.omit
              )
summary(lme_fr)
vif(lme_fr) # test VIF here, because I'm not sure how to do it in GAMM yet...? All low VIF (<3) when random = island, but when random = site, VIFs jump up (for scarid_den & pc2)
```

## GAMM
- for GAMMs - seems model results are the same whether random = site or island?

```{r,eval=F}
gamm<-gamm(fr~phase+s(length_cm)+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_id_pca, species_code=="qup"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm$gam)
plot(gamm$gam,pages=1)
```
```{r,eval=F}
gamm1<-gamm(fr_mean~phase+species+s(pc1_bent)+s(pc2_bent)+s(scar_bm)+s(scar_den), 
           family=gaussian(link=identity),
           data=filter(sum_ssp_pca, species_code!="rbp" & island!="Barbuda"),
           random=list(site=~1),
           method="REML"
           )
summary(gamm1$gam)
AIC(gamm1$lme)
plot(gamm1$gam,pages=1)

```

------------------------
To Do as of Nov. 7  
* boosted regression trees ecosphere 2017 adrians paper  
* species as random effect