### 2018-11-26 r/Models.Rmd

---
title: "Scarid behavior GAM/GAMMS"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

library(startR)
library(here)
library(tidyverse)
library(mgcv)
library(car)
library(ggfortify)

sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Examining predictor variables
Notes: in models using benthic-only PCA, scarid biomass and scarid density are highly collinear with benthic PCA variables (understandably). I don't think I can justify separating them, even though they are my ultimate variables of interest. I can use PCA loadings to interpret model outcomes. Carnivore biomass and spearfishing presence is also highly correlated with scarid biomass and thus benthic PCA variables - I am excluding it for now.

```{r,warning=FALSE,echo=FALSE}
predictors <- sum_site %>% 
  select(tot_bm,tot_den,scar_bm,scar_den,carn30_bm,carn30_den,
         cca_cover,lc_cover,ma_cover,ma_canopy,ta_cover,ta_canopy,rugosity,spear_daily) %>% 
  gather() %>% 
  na.omit()

ggplot(predictors,aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

```{r}
pca_all <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity, scar_bm, scar_den)),
  center = T,scale = T)
autoplot(pca_all, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "salmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()

pca_bent <- prcomp(na.omit(sum_site %>% 
              remove_rownames %>% 
              column_to_rownames(var="site") %>%
              # filter(island!="Barbuda") %>%
              select(ma_cover, ma_canopy, ta_cover, ta_canopy, rugosity)),
  center = T,scale = T)

sum_site <- sum_site %>% 
  # filter(island!="Barbuda") %>%
  bind_cols(as.data.frame(pca_bent$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_bent=PC1,
         pc2_bent=PC2) %>% 
  bind_cols(as.data.frame(pca_all$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>% 
  left_join(select(sum_site,site,pc1_bent,pc2_bent,pc1_all,pc2_all), by = "site")

```

```{r,echo=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
pairs(~scar_bm + pc1_bent + pc2_bent, data=sum_site,
      lower.panel=panel.smooth, upper.panel=panel.cor, 
      pch=20, main="Selected Fish & Benthic Community Variables", na.action = na.omit)
```

```{r}
m_data <- sum_id_pca %>% filter(species_code!="rbp")
vif(lm(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data))
vif(lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
        random=~1|island, 
        na.action = na.omit
        ))
```

Confirming I can't use scarid biomass and/or density as separate from benthic PCA variables. VIF of scar_bm is just slightly >3. Removing rugosity from benthic PCA reduces scar_bm ~ pc1 correlation slightly, but still >0.7.

BUT: VIF done on LME where random=site or random=island shows values <3 when using benthic PCA values + scarid biomass... *stay tuned...*

# Examining potential response variables
Feeding rate (fr) is primary variable of interest
```{r,echo=FALSE,warning=FALSE}
response <- sum_id %>% 
  select(g_frac,fr,br,for_dur,for_bites) %>% 
  gather() %>% 
  na.omit()

ggplot(response,aes(value)) + 
  facet_wrap(~ key, scales = "free") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") 
```

# Model as LME first, check model validity  
I won't get too far into dealing with heteroskedasticity here, because it may be improved when switching to GAMM
```{r}
m_data <- sum_id_pca %>% filter(species_code!="rbp" & fr<=3000) # removing outlier
lme1 <- lme(fr ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme1)
vif(lme1)
plot(lme1, which = c(1), col = 1)
plot(m_data$length_cm, resid(lme1))
plot(m_data$scar_bm, resid(lme1)) # highest scar_cm site is big jump from other intervals - scale?
plot(m_data$pc1_bent, resid(lme1))
plot(m_data$pc2_bent, resid(lme1)) # potential interval/scale issues?
plot(factor(m_data$phase), resid(lme1))
plot(factor(m_data$species), resid(lme1)) # much greater spread in S. vetula data than S. viride

```

_Dealing with heteroskedasticity_: clear heteroskedasticity in my data, likely with multiple sources...  
- differences in variance between species, but modelling species independently still does not fully account for heteroskedasticity (below)
- scar_cm and pc2_bent have unevenly distributed intervals - scale?

_Options/notes from Zuur et al. 2009 (Ch.4)_:  
- add variance covariate (p.78)
- log transform DV (p.131)

Dealing with species separately:

```{r}
# S. vetula only
qup_data <- sum_id_pca %>% filter(species_code=="qup")
lme2 <- lme(fr ~ phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=qup_data,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme2)
vif(lme2)
plot(lme2, which = c(1), col = 1)

# S. viride only
vir_data <- sum_id_pca %>% filter(species_code=="stop")
lme3 <- lme(fr ~ phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=vir_data,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme3)
vif(lme3)
plot(lme3, which = c(1), col = 1)
```

Log transforming feeding rate variable - still heteroskedasticity, largely from observations with fr=0?

```{r}
m_data_tr <- sum_id_pca %>% 
  filter(species_code!="rbp" & fr<=3000) # removing outlier
  # mutate(fr_c = scale(fr, scale = F)) # trying to center fr variable
lme4 <- lme(log10(fr/100 + 1) ~ species + phase + length_cm + scar_bm + pc1_bent + pc2_bent, data=m_data_tr,
    random = ~1 | island, 
    na.action = na.omit
    )
summary(lme4)
vif(lme4)
plot(lme4, which = c(1), col = 1)
```

Because of 0 values,

# Switching to gamm, checking model validity

Notes:  
- GLS: option for model when homogeneity of variances is not satisfied, can add weights for residuals. Keep reading Zuur et al. and decide if I need to go beyond this (e.g. GAMM)

```{r}
m1 <- gam(fr ~ factor(species) + s(length_cm, bs = "cs") + s(scar_bm, bs = "cs") + 
            s(pc1_bent, bs = "cs") + s(pc2_bent, bs = "cs"), data = m_data)
summary(m1)
plot(m1)
gam.check(m1)

e <- resid(m1)
coplot(e ~ length_cm | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc1_all | species, data = m_data) # differences in res. dist. between species
coplot(e ~ pc2_all | species, data = m_data) # inconsistent res.
coplot(e ~ scar_bm | species, data = m_data)

m2 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + 
          s(pc1_bent, bs = "cs") + s(pc2_bent, bs = "cs") +
          s(length_cm, by = as.numeric(species_code=="qup")), 
          data = m_data)
summary(m2)
gam.check(m2)

m3 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
          s(pc1_bent, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m3)
gam.check(m3)

m4 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
          s(pc2_bent, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m4)
gam.check(m4)

m5 <- gam(fr ~ factor(species_code) + s(length_cm) + s(scar_bm) + s(pc1_bent) + s(pc2_bent) +
          s(pc1_bent, by = as.numeric(species_code == "qup")) +
          s(pc2_bent, by = as.numeric(species_code == "qup")) +
          s(length_cm, by = as.numeric(species_code == "qup")) +
          s(scar_bm, by = as.numeric(species_code == "qup")),
          data = m_data)
summary(m5)
plot(m5)
gam.check(m5)
```

_Model validity notes_:  
- Violates assumption of homogeneity of variance (increasing variance as predictor values increase)
- Satisfies assumption of normality

*need to add interaction terms*
```{r}
gamm1 <- gamm(fr ~ species_code + phase + s(length_cm) + s(pc1_bent) + s(pc2_bent) + s(scar_bm) +
                s(length_cm, by = as.numeric(species_code == "qup")) +
                s(pc1_bent, by = as.numeric(species_code == "qup")) +
                s(scar_bm, by = as.numeric(species_code == "qup")), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           weights = varIdent(form =~ 1 | species),
           data = m_data
           )
summary(gamm1$gam)
plot(gamm1$gam,pages=1)
plot(gamm1$lme) # shows heteroskedasticity
AIC(gamm1$lme)
```

*Re: Nash et al. 2016: should I be aggregating data to the site level?*
```{r}
m_data2 <- sum_ssp_pca %>% filter(species_code!="rbp") # currently doesn't work when removing rbp data
gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m) + s(pc1_bent) + s(pc2_bent) + s(scar_bm), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = sum_ssp_pca
           )
summary(gamm2$gam)
plot(gamm2$gam,pages=1)
plot(gamm2$lme) # not too heteroskedastic?
AIC(gamm2$lme)

# Confirming low VIF among predictor variables with site-level model
lme1 <- lme(fr_mean ~ phase + length_m + scar_bm + pc1_bent + pc2_bent, data=sum_ssp_pca,
    random = ~1 | island, 
    na.action = na.omit
    )
vif(lme1)
```