### 2019-02-01 r/Analysis.Rmd
feb1 updates

---
title: "Figures + Analysis"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Outline of selected figures and analyses:

- Grazing differences between islands: (site or island level?)  
-- ANOVAs to test statistical significance of differences here  
- Total grazing impact of site-level biomass: measured vs. reported grazing rates (add line or color by feeding rate)  
- Something looking at relationship with algal index vs. scarid biomass 
- SEM?

Notes: 
- re: lms/gamms: have been stressed about collinearities in the data, but some aren't that bad... return to models once I have a better handle on what I want to use (and make sure I'm looking at mixed effects models only?)  
- Exclude Barbuda from this analysis...?  
- From Nash et al. 2016:  
  + used GAMMs with site and reef zone as random effects
  + aggregated data to the site level
  + used mean or median values based on skew of different predictor/response variables (within each site)

```{r}
# set up
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(startR)
library(here)
library(tidyverse)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
library(ggfortify)
library(ggsignif) # boxplot stats
library(EnvStats) # boxplot stats
library(car)
library(nlme)
library(gamm4)
```
```{r}
sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>%
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(length_cm>=15)
fish <- read.csv(here("data","fish.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Island - level differences in grazing rates

```{r}
# Violin plot of inter-island differences
ggplot(filter(sum_id, 
              phase=="i" & length_cm>=16 & length_cm<=24 & species_code!="rbp"), 
       aes(factor(island), fr)) + 
  geom_violin(aes(fill=island)) +
  geom_boxplot(width=0.1) + 
  labs(y = "Feeding Rate (bites/hr)", x = "", title = "Hourly feeding rates (i. phase, 16-24cm)") +
  geom_signif(comparisons = list(c("Antigua", "Barbuda"),c("Antigua","Bonaire"),c("Barbuda","Bonaire")), map_signif_level=TRUE) +
  facet_grid(.~species) + 
  stat_n_text() +
  scale_fill_brewer(palette="Blues") +
  theme_bw() +
  theme(strip.text = element_text(face = "italic")) +
  theme(plot.title = element_text(hjust = 0.5))

# Length-weight relationships between islands
ggplot(filter(sum_id, species_code != "rbp" & island != "Barbuda"), 
       aes(x = length_cm, y = fr, colour = island)) +
  facet_grid(.~species) +
  geom_point(shape = 18) +
  scale_shape(solid = FALSE) +
  geom_smooth(method = lm, fullrange = FALSE) +
  labs(y = "Feeding Rate (bites/hr)", x = "Length (cm)", title = "Hourly feeding rate") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_color_manual(values = c("navy", "darkseagreen", "slategray2"))

# Calculating length-weight coefficients

fr_anu_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "stop"))
summary(fr_anu_vir) # fr(Antigua, S. viride, initial) = 445-8*L

fr_anu_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Antigua" & phase == "i" & species_code == "qup"))
summary(fr_anu_vet) # fr(Antigua, S. vetula, initial) = 1030-12*L

fr_bon_vir <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "stop"))
summary(fr_bon_vir) # fr(Bonaire, S. viride, initial) = 1057-17*L

fr_bon_vet <- lm(fr~length_cm, data=filter(sum_id, island == "Bonaire" & phase == "i" & species_code == "qup"))
summary(fr_bon_vet) # fr(Bonaire, S. vetula, initial) = 1347+10*L
```

$$fr(Antigua, S. viride, i) = 445-8*L$$
$$fr(Antigua, S. vetula, i) = 1030-12*L$$
$$fr(Bonaire, S. viride, i) = 1057-17*L$$
$$fr(Bonaire, S. vetula, i) = 1347+10*L$$

# Comparison to published equations

## Predicted feeding rates and bite sizes from standard equations
Bite sizes as predicted by Bruggemann/Bozec:
$$bs(S. vetula) = 4.013*10^{-4}*L^2$$
$$bs(S. viride) = 5.257*10^{-4}*L^2$$

Feeding rates as published by Bruggemann/Bozec
$$fr(Published, S. viride, i) = 1089-17*L-56 $$
  *(extra term because phase modifier is not used here (i=1))*
$$fr(Published, S. vetula, initial) = 3329-33*L$$

## Adding predicted feeding rates from observed relationships

```{r}
fish_fr <- fish %>% 
  filter(species_code %in% c("stop", "qup") & island %in% c("Antigua","Bonaire")) %>%
  mutate(
    fr_pers = ifelse(island == "Antigua" & species_code == "stop", 445-8*length,
                     ifelse(island == "Antigua" & species_code == "qup", 1030-12*length,
                            ifelse(island == "Bonaire" & species_code == "stop", 
                                   1057-17*length, 1347+10*length))),
    fr_pub = ifelse(species_code == "stop", 1089-17*length-56, 3329-33*length),
    bs = ifelse(species_code == "stop", 5.257*10^-4*length^2, 4.013*10^-4*length^2),
    ag_pers = fr_pers*bs,
    ag_pub = fr_pub*bs
)

ag_site <- fish_fr %>% group_by(site,island) %>%
  summarize(personal=sum(ag_pers)/1000,published=sum(ag_pub)/1000)
ag_melt <- ag_site %>% as.data.frame() %>% 
  gather(source, area_grazed, -c(site,island))

ggplot(filter(ag_melt, island == "Antigua"),
      aes(x=reorder(site, -area_grazed), y=area_grazed)) +
        geom_bar(aes(fill = source), position = "dodge", stat = "identity") +
        scale_fill_manual(values = c("black", "slategray2")) +
        labs(x = "Site", 
             y = expression(Area~grazed~(m^2)~per~Hour), 
            title = expression(Predicted~cumulative~grazing~impact~on~Antiguan~reefs~(m^2/hr))) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggsave(here("figs/grazing_imp.png"))
```

# Bite content hypothesis

Fish in high algal sites may be taking fewer bites because they can consume more algae per bite. Let's try and add a rough estimate of bite content as a function of algal canopy height.

Notes/assumptions:
* Using turf algal canopy height (for now, no macroalgal component - though S. viride does eat some macrogalgae)
* Max bite height will be capped at radius of bite size ($\sqrt{BS}/\pi$)

```{r}
# grazing impact of fish from fish surveys
fish_grazed <- fish_fr %>% 
  left_join(select(sum_site, ta_canopy, site), by = "site") %>%
  mutate(
    bh_capped = ifelse(
      ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr_pers*bv
 )

# grazing impact of fish from behavioral surveys
sum_id_vg <- sum_id %>%
  filter(species_code != "rbp") %>%
  mutate(
    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr*bv,
    vg_capped = fr*bv_capped,
    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
) %>%
  arrange(island)

sum_ssp_vg <- sum_id_vg %>% 
  group_by(site, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    vg_mean=mean(vg),
    vg_se=sd(vg)/sqrt(n),
    vg_capped_mean=mean(vg_capped),
    vg_capped_se=sd(vg_capped)/sqrt(n)
    ) %>% 
  select(-n)

ggplot(filter(sum_ssp_grazed, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))

ggplot(filter(sum_ssp_grazed, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_capped_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_capped_mean-vg_capped_se, ymax=vg_capped_mean+vg_capped_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (capped height, initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```