### 2019-03-01 r/Ch2 outline_V3.Rmd
updates

---
title: "Variance in scarid grazing behavior across Caribbean reef sites"
author: "Molly Wilson"
date: "3/1/2019"
output: html_document
---

# Variance in scarid grazing behavior across Caribbean reef sites
### Formatted as: Coral Reefs note (<2900 words)

```{r, echo=F, message=F, warning=F}
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggfortify)
library(mgcv)
library(MuMIn)
library(knitr)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

## Introduction

Herbivorous fish play a critical role in coral reef ecosystems by supressing algae and facilitating coral growth, recruitment, and survival (). In recent decades, coral reefs have become increasingly threatened by anthropogenic impacts including the fishing down of herbivorous fish populations (). Increasing attention has been given to understanding how much herbivory is needed to sustain coral reef health and to identify thresholds of herbivore biomass that can be used to guide fisheries management targets (Karr et al. 2015, McClanahan et al. 2011)

However, not all herbivores are the same in terms of their grazing impact on reefs. Recent studies have highlighted functional differences across herbivorous fish species in terms of their feeding selectivity, feeding morphology, etc. (e.g. Adam et al. 2018) [**add more here**]

In addition to these functional differences found across species, herbivore behavior may also drive variation in herbivore function across different reef environments. Empirical studies have demonstrated the sensitivity of herbivore behavior to predator presence (), reef rugosity (), grazing behaviors of proximate herbivores (Gil et al. 2018), and algal abundance, distribution, and nutritional content (). 
[**add more here**]
[These behavioral variations have not been linked to larger-scale variations - except Madin's seascape papers...]
[Maybe even highlight Madin papers - these behavioral differences have the potential to drive 
Madin et al.'s study showing how predator abundance drove herbivore feeding ranges/grazing halos]

Despite a growing understanding of the sensitivity and ecological importance of herbivore behavior, current efforts to predict herbivore impact on reefs assume species-specific feeding rates that vary only with individual fish size. There is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales. Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications. We demonstrate a need for further experimental and modelling work to unpack these drivers and determine their ecological consequences. We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems.

# Methods

*Data collection*
In March through August of 2017, I collected data from 13 reef sites in Bonaire, Antigua, and Barbuda (Fig. 1). At each site I conducted fish, benthic, and rugosity surveys to quantify reef (condition). I used a modified AGRRA protocol with 30m by 4m belt transects and 10m point-count transects for fish and benthic survets, respectively. To assess rugosity I measured the length of reef contour under each meter of a taut 10m leadline (sensu ___). I carried out a minimum of X fish transects and X benthic and rugosity transects per site.

At each site, I observed grazing behavior of two dominant scarid species, *Sparisoma viride* and *Scarus vetula*. I targetted these species because of their relative abundance as well as their contrasting grazing mechanisms. *S. viride* is classified as an excavator, taking [_] (), while *S. vetula* is a grazer, taking shallower bites from [_] (). I quantified grazing behavior during two minute follows, 
* Observational behavioral data:
    - Species: , initial and terminal phase greater than 15cm, focus on only initial phase here
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables ([table]): feeding rate, bite rate, feeding intensity

```{r}
Variable <- c('Time spent grazing','Bite rate','Feeding rate','Feeding intensity')
Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
Unit <- c('fraction','bites/second','bites/minute','bites/foray')
var.table <- data.frame(Variable,Definition,Unit)
kable(var.table)
```

*Analysis*

* Fish biomass using length-weight conversions (appendix)
* Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
* Biplots w predictors - loess smoothers (switch to something else?)
* Bite content: calculated proxy for bite content by multiplying estimated bite sizes (based on species and length of each individual fish surveyed, equations from Bruggemann/Mumby) by site level averages of turf algal canopy height to account for potential differences in amount of algae consumed in low vs. high canopy sites

## Results and discussion

## Variation among sites in multiple components of feeding behavior
Fig. X shows intraspecific variation in both feeding rate (bites per hour) and feeding intensity (bites per foray) among sites. (Bite rates while grazing were relatively constant, indicating that differences in overall feeding rates were driven by differences in the amount of time spent grazing).

```{r, eval = F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean feeding rates of individuals (initial phase)", 
         y="Feeding rate (bites/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

Notes:

* Can remove color coding by island if this is distracting. Unsure whether I should be highlighting differences between islands or if I should just be highlighting variation among sites (mixed feedback on this)

```{r, echo = F, message = F, warning = F}
# Panel graph of site differences

fr_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Feeding rate \n (bites/min)", 
       title = "S. vetula") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

fr_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       title = "S. viride") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

br_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Bite rate \n(bites/sec)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

br_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

fb_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Feeding intensity\n(bites/foray)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

fb_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

gfrac_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Time grazing \n(fraction)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        axis.title.y = element_text(size=10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

gfrac_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

legend_plot <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() +
  theme(legend.position = "right",
        legend.title = element_blank())

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  }
legend <- get_legend(legend_plot)

grid.arrange(
  arrangeGrob(fr_qup, fr_stop, fb_qup, fb_stop, gfrac_qup, gfrac_stop, ncol=2, heights = c(1,1,1.5)), 
              arrangeGrob(legend), 
              ncol=2, widths = c(5,1))

```

## Several potential underlying mechanisms

```{r, echo=F, message=F, warning=F}
fr_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = expression(atop(NA, atop(textstyle('Scarid biomass'),
                               textstyle('(kg/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.title.x = element_text(size = 9)
        )

fr_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Conspec. density'),
                               textstyle('(indv/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Rugosity'),
                               textstyle('  ')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('EAM canopy'),
                               textstyle('height (mm)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('EAM cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Macroalgal cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 9),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

label_vet = textGrob("S. vetula", gp=gpar(fontface="italic"), rot=270)
label_vir = textGrob("S. viride", gp=gpar(fontface="italic"), rot=270)

grid.arrange(left = "Feeding rate (bites/min.)",
  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
  nrow=2, heights = c(1,1.2)) 
```

While we do not seek to conclusively establish mechanisms here, we discuss potential underlying mechanisms that could be driving these trends in herbivore feeding behaviors

* Bite content/nutrition hypothesis: scarids in higher canopy sites are consuming more per bite, thus taking fewer bites - implications for spatial grazing impact. Previous work showing that scarids vary grazing with nutrient availability/distribution (Carlson, Tootell, Mumby, etc.)

```{r, echo=F, message=F, warning=F}
# grazing impact (vg ~ volume grazed) of fish from behavioral surveys
sum_id_vg <- sum_id %>%
  filter(species_code != "rbp") %>%
  mutate(
    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr*bv,
    vg_capped = fr*bv_capped,
    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
) %>%
  arrange(island)

# summing by site, species, and phase
sum_ssp_vg <- sum_id_vg %>% 
  group_by(site, ta_canopy, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    vg_mean=mean(vg),
    vg_se=sd(vg)/sqrt(n),
    vg_capped_mean=mean(vg_capped),
    vg_capped_se=sd(vg_capped)/sqrt(n),
    fr_mean=mean(fr)
    ) %>% 
  select(-n)

vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.title = element_blank())

vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
       x = "", 
       title = "S. vetula") +
  theme_bw() + 
  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
        axis.title.y = element_text(size=10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "italic")
        )

vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +    
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "S. viride") +
  theme_bw() + 
  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "italic")
        )

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  }
legend <- get_legend(vg_facet)

grid.arrange(
  arrangeGrob(vg_qup, vg_stop, ncol=2), 
              arrangeGrob(legend), 
              ncol=2, widths = c(5,1), heights = c(1))

ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
  geom_point(shape=18) +
  facet_grid(~species) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        )
```

* Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)
* Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e. those sampled here). (Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.

## Ecological implications

* Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid
    * Feeding intensity: less intense feeding/more distributed bites in degraded reefs may also reduce actual substrate made available to recruiting corals (~ grazing concentration lit. here)

## Management implications + future research needs

* Limitations to this study: all these potential drivers/mechanisms are understandably correlated in observational work
    * Room for experimental work to unpack/distinguish drivers
* Need for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions. Current paradigm of constant species-level grazing behaviors may be inaccurate. In Caribbean case, most models use Bonaire data which may not be representative of dynamics on more degraded Caribbean reefs.
* Need for framework to help incorporate behavioral effects into ecosystem management
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health

# Potential additions:

## PCA
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
                     scar_bm, carn30_bm)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_brewer(palette = "Blues") +
    theme_bw() +
    theme(legend.title = element_blank())
ggsave(here("figs/pca_pred.png"))

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```

## GAMM results

### Predicting feeding rate:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Species + Phase + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Phase + Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Phase + Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Species + Phase + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Species + Phase + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g10 <- c("g10", "Species + Phase + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g11 <- c("g11", "Phase + Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

```{r, echo=F, message=F}
kable(gtab_fr[1:4, ])
plot(gamm3$gam,pages=1)
```

### Predicting feeding intensity:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

fi_gamm1 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
# summary(fi_gamm1$gam)
# plot(fi_gamm1$gam,pages=1)
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(fi_gamm1$lme), summary(fi_gamm1$gam)$r.sq)

fi_gamm2 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(fi_gamm2$lme), summary(fi_gamm2$gam)$r.sq)

fi_gamm3 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species", AICc(fi_gamm3$lme), summary(fi_gamm3$gam)$r.sq)

fi_gamm4 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(fi_gamm4$lme), summary(fi_gamm4$gam)$r.sq)

fi_gamm5 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Phase + Size~Species + PC1~Species", AICc(fi_gamm5$lme), summary(fi_gamm5$gam)$r.sq)

fi_gamm6 <- gamm(for_bites_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Species + Phase + Size + PC1~Species", AICc(fi_gamm6$lme), summary(fi_gamm6$gam)$r.sq)

fi_gamm7 <- gamm(for_bites_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Species + Phase + Size~Species + PC1", AICc(fi_gamm7$lme), summary(fi_gamm7$gam)$r.sq)

fi_gamm8 <- gamm(for_bites_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Phase + Size + PC1~Species", AICc(fi_gamm8$lme), summary(fi_gamm8$gam)$r.sq)

fi_gamm9 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Phase + Size~Species + PC1", AICc(fi_gamm9$lme), summary(fi_gamm9$gam)$r.sq)

gtab_fi <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9)) %>% setNames(c("g", "Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

```{r, echo=F, message=F}
kable(gtab_fi[1:3, ])
plot(gamm6$gam,pages=1)
```