### 2019-03-02 r/Ch2 outline_V3.Rmd
organizing/updates to v3

---
title: "Variance in scarid grazing behavior across Caribbean reef sites"
author: "Molly Wilson"
date: "3/1/2019"
output: html_document
---

# Variance in scarid grazing behavior across Caribbean reef sites
### Formatted as: Coral Reefs note (<2900 words)

```{r, echo=F, message=F, warning=F}
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(grid)
library(ggfortify)
library(mgcv)
library(MuMIn)
library(knitr)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island) %>%
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  mutate(island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))) %>%
  arrange(island)
```

**Abstract:** Herbivorous fish increase coral growth and survival by grazing down algal competitors. As coral reef health continues to decline globally, maximizing herbivory on reefs has become a primary management goal. However, existing management targets assume consistent grazing behavior across different reef systems. Here I document the grazing behavior of two scarid species at 13 sites in Antigua, Barbuda, and Bonaire. My analyses show significant differences in overall feeding rates, time spent grazing, and intensity of grazing across reef sites, which may alter the ecological impact of a given scarid population. I propose several potential mechanisms for these grazing differences, including variations in bite content which may make bite rates a misleading measure of overall grazing impact. As we set management targets to enhance herbivory on coral reefs, it is critical that we are accurately estimating the grazing impact of scarid populations. This paper demonstrates the variability of grazing behavior across different reef sites and calls for further investigation of the drivers and ecological implications of these inconsistencies.

## Introduction

Herbivorous fish play a critical role in coral reef ecosystems by supressing algae and facilitating coral growth, recruitment, and survival (). In recent decades, coral reefs have become increasingly threatened by anthropogenic impacts including the fishing down of herbivorous fish populations (). Increasing attention has been given to understanding how much herbivory is needed to sustain coral reef health and to identify thresholds of herbivore biomass that can be used to guide fisheries management targets (Karr et al. 2015, McClanahan et al. 2011)

However, not all herbivores are the same in terms of their grazing impact on reefs. Recent studies have highlighted functional differences across herbivorous fish species in terms of their feeding selectivity, feeding morphology, etc. (e.g. Adam et al. 2018) [**add more here**]

In addition to these functional differences found across species, herbivore behavior may also drive variation in herbivore function across different reef environments. Empirical studies have demonstrated the sensitivity of herbivore behavior to predator presence (), reef rugosity (), grazing behaviors of proximate herbivores (Gil et al. 2018), and algal abundance, distribution, and nutritional content (). 
[**add more here**]
[These behavioral variations have not been linked to larger-scale variations - except Madin's seascape papers...]
[Maybe even highlight Madin papers - these behavioral differences have the potential to drive 
Madin et al.'s study showing how predator abundance drove herbivore feeding ranges/grazing halos]

Despite a growing understanding of the sensitivity and ecological importance of herbivore behavior, current efforts to predict herbivore impact on reefs assume species-specific feeding rates that vary only with individual fish size. There is very limited data on how herbivore feeding behaviors might vary across reef systems and whether we can accurately assume constant species roles at regional scales. Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and discuss potential underlying mechanisms, ecological consequences, and management implications. We demonstrate a need for further experimental and modelling work to unpack these drivers and determine their ecological consequences. We urge for the incorporation of behavioral effects in coral reef herbivore management and caution against assuming constant species function across reef systems.

# Methods

*Data collection*
In March through August of 2017, I collected data from 13 reef sites in Bonaire, Antigua, and Barbuda (Fig. 1). At each site I conducted fish, benthic, and rugosity surveys to quantify reef (condition). I used a modified AGRRA protocol with 30m by 4m belt transects and 10m point-count transects for fish and benthic survets, respectively. To assess rugosity I measured the length of reef contour under each meter of a taut 10m leadline (sensu ___). I carried out a minimum of X fish transects and X benthic and rugosity transects per site.

At each site, I observed grazing behavior of two dominant scarid species, *Sparisoma viride* and *Scarus vetula*. I targetted these species because of their relative abundance as well as their contrasting grazing mechanisms. *S. viride* is classified as an excavator, taking [_] (), while *S. vetula* is a grazer, taking shallower bites from [_] (). I quantified grazing behavior during two minute follows, 
* Observational behavioral data:
    - Species: , initial and terminal phase greater than 15cm, focus on only initial phase here
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables ([table]): feeding rate, bite rate, feeding intensity

```{r}
Variable <- c('Time spent grazing','Bite rate','Feeding rate','Grazing concentration')
Definition <- c('Proportion of time spent grazing during follow','Bite rate while actively feeding','Bite rate over total follow time','Number of consecutive bites taken during single grazing foray')
Unit <- c('fraction','bites/second','bites/minute','bites/foray')
var.table <- data.frame(Variable,Definition,Unit)
kable(var.table)
```

*Analysis*

* Fish biomass using length-weight conversions (appendix)
* Site-level behavioral summary stats taken from subset of individuals based on size to avoid bias
* Biplots w predictors - loess smoothers (switch to something else?)
* Bite content: calculated proxy for bite content by multiplying estimated bite sizes (based on species and length of each individual fish surveyed, equations from Bruggemann/Mumby) by site level averages of turf algal canopy height to account for potential differences in amount of algae consumed in low vs. high canopy sites

## Results and discussion

Feeding rates, time spent grazing, and grazing concentration varied significantly across sites for both *S. vetula* and *S. viride* populations (Fig. 2). [Fig. 2A also shows substantially lower feeding rates in most Antiguan (*S. vetula* and *S. viride*) and Barbudan (*S. vetula*) sites as compared standard values used in Caribbean herbivory models - *if I include reference lines...*]. Differences in site-level feeding rates stemmed from differences in the fraction of time that fish speng grazing (Fig. 2B) as opposed to differences in bite rates while activitely feeding, which remained relatively constant accross sites (appendix).

These behavioral variations may have important ecological consequences and mangement implications. If we use grazing data from one area to predict the impact of a given species in other parts of the region, we may be overlooking important behavioral differences and misrepresenting grazing levels of different scarid populations. Feeding rates are used in herbivory models to calculate the area grazed by a given fish community. In the Caribbean case, most standard feeding rate equations are obtained from data collected in Bonaire. My results suggest that applying Bonaire feeding rates to other Caribbean reef systems may overestimate grazing impact.

Grazing concentration also plays a key role in determining the ecological impact of coral reef herbivores. More concentrated grazing may be more effective in removing algae and clearing substrate for growing or newly recruiting corals than bites dispersed throughout the reefscape. [**add citations here**]

```{r, eval = F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean feeding rates of individuals (initial phase)", 
         y="Feeding rate (bites/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 90, hjust = 1), 
          plot.title = element_text(hjust = 0.5))
```

```{r, echo = F, message = F, warning = F}
# Panel graph of site differences

ref_qup15 <- (3329-33*15)/60 # published feeding rate (bites/min)
ref_qup30 <- (3329-33*30)/60
ref_stop15 <- (1089-17*15-56)/60
ref_stop30 <- (1089-17*30-56)/60

fr_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  #geom_abline(slope=0, intercept=ref_qup15,  col = "red",lty=2) +
  #geom_abline(slope=0, intercept=ref_qup25,  col = "red",lty=2) +
  #ylim(0, 50) +
  labs(y = "Feeding rate \n (bites/min)", 
       title = "S. vetula") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

fr_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=fr_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=fr_mean-fr_se, ymax=fr_mean+fr_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  #geom_abline(slope=0, intercept=ref_stop15,  col = "red",lty=2) +
  #geom_abline(slope=0, intercept=ref_stop25,  col = "red",lty=2) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       title = "S. viride") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, face = "italic"),
        plot.margin = margin(5, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

br_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Bite rate \n(bites/sec)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

br_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=br_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=br_mean-br_se, ymax=br_mean+br_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/hr)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

gfrac_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Time grazing \n(fraction)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=10),
        legend.position = "none")

gfrac_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=g_frac_mean-g_frac_se, ymax=g_frac_mean+g_frac_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        strip.text = element_text(face = "italic"), 
        axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

fi_qup <- 
  ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "Grazing concentration\n(bites/foray)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 0, 0, 5, unit = "pt"),
        axis.title.y = element_text(size=10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

fi_stop <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=for_bites_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  geom_errorbar(aes(ymin=for_bites_mean-for_bites_se, ymax=for_bites_mean+for_bites_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.margin = margin(0, 4, 0, 0, unit = "pt"),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none")

legend_plot <- 
  ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), 
         aes(x=factor(site, levels = sum_site$site), y=g_frac_mean)) +
  geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "") +
  theme_bw() +
  theme(legend.position = "right",
        legend.title = element_blank())

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  }
legend <- get_legend(legend_plot)

grid.arrange(
  arrangeGrob(fr_qup, fr_stop, gfrac_qup, gfrac_stop, fi_qup, fi_stop, ncol=2, heights = c(1,1,1.5)), 
              arrangeGrob(legend), 
              ncol=2, widths = c(5,1))

```

The variations in grazing behaviors demonstrated here may be driven by several potential underlying mechanisms. Figure 3 shows the relationships between various reef characteristics and scarid feeding rate. Because study sites expectedly varied across a number of interrelated fish and benthic dimensions (PCA - appendix?), it is impossible to conclusively establish any explicit mechanisms here. Instead I present several hypotheses that may be driving these observed behavioral trends and discuss evidence for and implications of each.

```{r, echo=F, message=F, warning=F}
fr_q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank()
        )

fr_s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = expression(atop(NA, atop(textstyle('Scarid biomass'),
                               textstyle('(kg/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"),
        axis.title.x = element_text(size = 9)
        )

fr_s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Conspec. density'),
                               textstyle('(indv/100m'^2*')')))),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Rugosity'),
                               textstyle('  ')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('EAM canopy'),
                               textstyle('height (mm)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('EAM cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 9),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

fr_s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean)) +
  geom_smooth(method="loess", span=1) +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = expression(atop(NA, atop(textstyle('Macroalgal cover'),
                               textstyle('(%)')))), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 9),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

label_vet = textGrob("S. vetula", gp=gpar(fontface="italic"), rot=270)
label_vir = textGrob("S. viride", gp=gpar(fontface="italic"), rot=270)

grid.arrange(left = "Feeding rate (bites/min.)",
  arrangeGrob(fr_q1, fr_q3, fr_q4, fr_q5, fr_q7, ncol=5, right = label_vet),
  arrangeGrob(fr_s1, fr_s3, fr_s4, fr_s5, fr_s7, ncol=5, right = label_vir), 
  nrow=2, heights = c(1,1.2)) 
```

*Bite content hypothesis.* Differences in scarid grazing preferences can be driven by (local differences in) nutrient availability and distribution (Carlson, Tootell, Mumby, etc.). Algal communities on Caribbean reefs can vary greatly in terms of canopy height, percent cover, species composition, and nutrient content ().

(no conclusive trends, but FR tended to decrease with increasing EAM canopy height and to some extent EAM and macroalgal percent cover)

In higher canopy sites, scarids may be obtaining more biomass of food per bite than in areas with heavily cropped algae, requiring them to take fewer bites to obtain the same nutritional intake.

Implications: 
-maybe better metric than bite/feeding rate
-reduced grazing impact if not clearing substrate
-reinforcing loop

```{r, echo=F, message=F, warning=F}
# grazing impact (vg ~ volume grazed) of fish from behavioral surveys
sum_id_vg <- sum_id %>%
  filter(species_code != "rbp") %>%
  mutate(
    bs = ifelse(species_code == "stop", 5.257*10^-4*length_cm^2, 4.013*10^-4*length_cm^2), # adding bite size equations from Mumby
    bh_capped = ifelse(ta_canopy < 2*sqrt(bs)/(22/7), ta_canopy, 2*sqrt(bs)/(22/7)), # if ta_canopy exceeds diameter of bite size, bite height is capped at bite size
    bv = bs*ta_canopy,
    bv_capped = bs*bh_capped,
    vg = fr*bv,
    vg_capped = fr*bv_capped,
    island = factor(island, levels = c("Bonaire","Antigua","Barbuda"))
) %>%
  arrange(island)

# summing by site, species, and phase
sum_ssp_vg <- sum_id_vg %>% 
  group_by(site, ta_canopy, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    vg_mean=mean(vg),
    vg_se=sd(vg)/sqrt(n),
    vg_capped_mean=mean(vg_capped),
    vg_capped_se=sd(vg_capped)/sqrt(n),
    fr_mean=mean(fr)
    ) %>% 
  select(-n)

vg_facet <- ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
    facet_grid(~species) +
    scale_fill_brewer(palette = "Blues") +
    labs(title="Mean grazing impact by volume (initial phase)", 
         y="Grazing impact (cm3/hr)", 
         x="Site") +
    theme_bw() +
    theme(strip.text = element_text(face = "italic"), 
          axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.title = element_blank())

vg_qup <- ggplot(filter(sum_ssp_vg, species_code == "qup" & phase == "i"), 
                 aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = expression('Grazing impact (cm'^3*'/hr)'), 
       x = "", 
       title = "S. vetula") +
  theme_bw() + 
  theme(plot.margin = margin(1, 0, 0, 5, unit = "pt"),
        axis.title.y = element_text(size=10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "italic")
        )

vg_stop <- ggplot(filter(sum_ssp_vg, species_code == "stop" & phase == "i"), 
                  aes(x=factor(site, levels = sum_site$site), y=vg_mean)) +
    geom_bar(stat = "identity", aes(fill = island), position = "dodge") +    
    geom_errorbar(aes(ymin=vg_mean-vg_se, ymax=vg_mean+vg_se), width=.2,
                 position=position_dodge(.9)) +
  scale_fill_brewer(palette = "Blues") +
  labs(y = "",
       x = "", 
       title = "S. viride") +
  theme_bw() + 
  theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
        axis.text.x = element_text(angle = 45, hjust = 1, size=7),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "italic")
        )

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
  }
legend <- get_legend(vg_facet)

grid.arrange(
  arrangeGrob(vg_qup, vg_stop, ncol=2), 
              arrangeGrob(legend), 
              ncol=2, widths = c(5,1), heights = c(1))

ggplot(filter(sum_ssp_vg, species_code != "rbp" & phase == "i"), aes(x = fr_mean, y = vg_mean)) +
  geom_point(shape=18) +
  facet_grid(~species) +
  labs(y = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        )
```

*Social feeding hypothesis.* more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory (e.g. Gil)

*Predation risk hypothesis.* Previous work has documented the suppressive effect of predator presence on herbivore feeding rates (Madin, Catano, McCauley?). 
- These studies have focused on natural predators (decoys or piscivore populations)
- While I found no clear relationship between large (>30cm) piscivore biomass and feeding trends (appendix), risk of predation from spearfishers may play a larger role in determining scarid grazing behavior at the reefs studied here. All Antiguan and Barbudan sites experience some degree of spearfishing pressure, while spearfishing has been banned in Bonaire since the 1970s. The 15-30cm individuals used in my behavioral observations are approaching size refuge from most natural predators, but would be likely targets for spearfishers. Both *S. vetula* and *S. viride* are commonly harvested species in Antigua and Barbuda.

- like with the social feeding hypothesis, spearfishing may have double impact of reducing fish and reducing grazing impact of remaining fish

- often interaction of rugosity (e.g. Madin, stay near structure)
(Incorporate rugosity here - Rogers, Madin, Catano). Did not find strong evidence for this here, but carnivore biomass was strongly correllated with herbivore biomass and rugosity.

## Ecological implications

* Potential reinforcing feedback loops in degraded reefs if high algal/low scarid reefs also have less grazing impact per scarid

## Management implications + future research needs

* Limitations to this study: all these potential drivers/mechanisms are understandably correlated in observational work
    * Room for experimental work to unpack/distinguish drivers
* Need for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions. Current paradigm of constant species-level grazing behaviors may be inaccurate. In Caribbean case, most models use Bonaire data which may not be representative of dynamics on more degraded Caribbean reefs.
* Need for framework to help incorporate behavioral effects into ecosystem management
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health

# Potential additions:

## PCA
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
                     scar_bm, carn30_bm)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_brewer(palette = "Blues") +
    theme_bw() +
    theme(legend.title = element_blank())
ggsave(here("figs/pca_pred.png"))

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```

## GAMM results

### Predicting feeding rate:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

gamm1 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Species + Phase + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Phase + Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Phase + Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Species + Phase + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Species + Phase + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + phase + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g10 <- c("g10", "Species + Phase + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g11 <- c("g11", "Phase + Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

```{r, echo=F, message=F}
kable(gtab_fr[1:4, ])
plot(gamm3$gam,pages=1)
```

### Predicting feeding intensity:

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")

fi_gamm1 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
# summary(fi_gamm1$gam)
# plot(fi_gamm1$gam,pages=1)
g1 <- c("g1", "Species + Phase + Size + PC1 + PC2", AICc(fi_gamm1$lme), summary(fi_gamm1$gam)$r.sq)

fi_gamm2 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Phase + Size + PC1", AICc(fi_gamm2$lme), summary(fi_gamm2$gam)$r.sq)

fi_gamm3 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Phase + Size~Species + PC1~Species", AICc(fi_gamm3$lme), summary(fi_gamm3$gam)$r.sq)

fi_gamm4 <- gamm(for_bites_mean ~ species_code + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4) + s(pc2_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Phase + Size~Species + PC1~Species + PC2", AICc(fi_gamm4$lme), summary(fi_gamm4$gam)$r.sq)

fi_gamm5 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Phase + Size~Species + PC1~Species", AICc(fi_gamm5$lme), summary(fi_gamm5$gam)$r.sq)

fi_gamm6 <- gamm(for_bites_mean ~ species + phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Species + Phase + Size + PC1~Species", AICc(fi_gamm6$lme), summary(fi_gamm6$gam)$r.sq)

fi_gamm7 <- gamm(for_bites_mean ~ species + phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Species + Phase + Size~Species + PC1", AICc(fi_gamm7$lme), summary(fi_gamm7$gam)$r.sq)

fi_gamm8 <- gamm(for_bites_mean ~ phase + s(length_m, k=4) + s(pc1_all, by = factor(species_code), k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Phase + Size + PC1~Species", AICc(fi_gamm8$lme), summary(fi_gamm8$gam)$r.sq)

fi_gamm9 <- gamm(for_bites_mean ~ phase + s(length_m, by = factor(species_code), k=4) + s(pc1_all, k=4), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Phase + Size~Species + PC1", AICc(fi_gamm9$lme), summary(fi_gamm9$gam)$r.sq)

gtab_fi <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9)) %>% setNames(c("g", "Model", "AICc", "Adj. R2")) %>% arrange(AICc)
```

```{r, echo=F, message=F}
kable(gtab_fi[1:3, ])
plot(gamm6$gam,pages=1)
```