### 2019-04-16 r/FID.Rmd
4.15

---
title: "Scarid Flight Response, Antigua"
author: "Molly Wilson"
date: "April 12, 2017"
output: html_document
---

```{r}
# prep
library(lazyeval)
library(here)
library(tidyverse)
library(reshape2)
```
```{r}
# import, format
fid <- read.csv(here("data","fid.csv"), stringsAsFactors=F) %>% 
  select(-c(FID.ft,FID.in,Bump.distance.ft,Bump.distance.in)) %>%
  janitor::clean_names()
```

# FID by species
```{r}
fid.sum <- fid %>%
  group_by(species, phase) %>%
  summarize(fid_mean = mean(fid_cm/100),
            fid_max = max(fid_cm/100)) %>%
  rename(max = fid_max,
         mean = fid_mean) %>%
  melt() %>%
  rename(fid = value)

ggplot(fid, aes(x = species, y = fid_cm)) +
  geom_boxplot()

ggplot(fid.sum %>% filter(phase == "Initial"), aes(x = species, y = fid, group = variable)) +
  geom_point(aes(color = variable))
```

# FID by spear, site
```{r}
lm.fid <-lm(data=fid, fid_cm ~ spear + site + rugosity + pre_approach_activity)
summary(lm.fid)

ggplot(fid, aes(x = pre_approach_activity,y=fid_cm)) + geom_point()
```

# bump distance by spear, site
```{r}
bump <- FID %>% filter(Bump.distance.cm != 0)
lmBump <-lm(data=bump,Bump.distance.cm~Spear + Site + Species + Rugosity + Pre.Approach.Activity)
summary(lmBump)
aov()

ggplot(bump, aes(x=Spear,y=Bump.distance.cm)) + geom_point(aes(color=Species))
```