### 2019-04-16 r/Follows.Rmd

---
title: "Antigua-Barbuda Fish Follows August 2017"
author: "Molly Wilson"
date: "8/8/2017"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

Wrangling of fish follow data from Antigua, Barbuda, and Bonaire, as well as integration of site-level data on fish, benthic, and rugosity (from Sites.rmd).

Output:
* "sum_id.csv": individual fish-level data (behavioral and survey data)
* "sum_site.csv": site-level summary data (behavioral and survey data)

```{r,message=FALSE,warning=FALSE}
# set up
library(startR)
library(tidyverse)
library(here)
```

```{r, warning=FALSE, message= FALSE}
# import data
site <- read.csv(here("data", "site.csv"), strip.white = T, stringsAsFactors = F) %>% select(-X)
sp_site <- read.csv(here("data","sp_site.csv"), stringsAsFactors = F) %>%
  filter(species %in% c("Sparisoma viride","Scarus vetula","Sparisoma aurofrenatum")) %>%
  select(-X,-family) %>%
  rename_at(vars(3:8), ~ paste0("consp_",.))

follows <- read.csv(here("data", "follows.csv"), strip.white = T, stringsAsFactors = F) %>%
  janitor::clean_names(case = "old_janitor") %>%
  filter(date != "" & date != "3/26/17") %>% 
  select(-activity_code) %>% 
  unite("id",c(site,diver,date,time,follow_number),sep="_", remove=FALSE) %>%
  mutate(
    time_sec = as.numeric(time_sec),
    time_frac = as.numeric(time_fraction),
    length_cm = as.numeric(length_cm),
    weight_g=as.numeric(weight_g),
    bites = as.numeric(ifelse(is.na(bites),0,bites))
  ) %>%
  replace(. =="#N/A", NA) %>%
  filter(length_cm >= 15)

# create df with single row for all fish followed
sum_id <- follows %>% 
  group_by(id, date, island, site, species, species_code, phase, length_cm, weight_g, follow_sec) %>% 
  summarize()
```

# Grazing behavior

## Time budgets 
Here I create a dataframe of time spent on each activity, per fish, in seconds and time fraction
*(could re-do with dcast____)*
```{r,warning=FALSE, message= FALSE}
tb <- follows %>% 
  group_by(id, activity, species, phase, length_cm) %>% 
  summarize(
    time = sum(time_sec), 
    time_frac = sum(time_frac))

# Add time budgets per activity for each fish to 'sum_id'

tbgraze <- tb %>% 
  filter(activity == "graze") %>% 
  subset(select=c(id,time,time_frac)) %>% 
  rename(
    g_sec=time,
    g_frac=time_frac) # keep seconds for calculating grazing rates
tbclean <- tb %>% 
  filter(activity == "cleaned") %>% 
  subset(select=c(id,time_frac)) %>% 
  rename(clean_frac=time_frac)

# join fish-level data to sum_id dataframe
sum_id <- sum_id %>% 
  left_join(tbgraze, by= "id") %>% 
  left_join(tbclean, by = "id") # %>%
  # replace(is.na(.),0) %>%
  # mutate(length_cm = ifelse(0,NA,length_cm)
```

## Grazing rates
```{r, warning=FALSE, message= FALSE}
# bites/second during total follow time and during grazing forays (including G1 instances)
bites <- follows %>% 
  group_by(id) %>% 
  summarise(bites = sum(bites))
sum_id <- sum_id %>% 
  left_join(bites, by="id") %>%
  mutate(
    br=bites/g_sec,
    fr=bites/follow_sec*60,
    tdb=bites/follow_sec*60*60*10.5
    ) %>%
  replace(is.na(.),0)

# # Bites/second during grazing forays, excluding G1 instances
# bites_noG1 <- follows %>% filter(activity=="graze"&Bites>1&Time_sec>1) %>% group_by(id) %>% summarize(Bites_noG1 = sum(Bites), g_sec_noG1=sum(Time_sec)) 
# sum_id <- left_join(sum_id, bites_noG1, by="id") %>% mutate(br_g_noG1=Bites_noG1/g_sec_noG1)
# sum_id$br_g_noG1[is.na(sum_id$br_g_noG1)] <- 0
```

## Foray length and G1 bites
*Should I remove incomplete forays? What about forays >2min?*
```{r,warning=FALSE, message= FALSE}
forays <- follows %>% filter(activity == "graze")
forays_c <- follows %>% # currently don't use this, but might want to only use complete forays
  filter(
    !(activity == "graze" & min == 0) | # filters forays starting at time=0
    !(lead(activity) == "end") # filters forays overlapping with end of follow
  )
sum_forays <- forays %>% 
  group_by(id) %>% 
  summarize(for_dur=mean(time_sec),for_bites=mean(bites))

sum_id <- sum_id %>% left_join(sum_forays, by= "id")

# # Frequency of G1 forays
# G1 <- forays %>% filter(Bites==1&Time_sec==1) %>% group_by(id) %>% dplyr::summarize(G1_freq=n())
# sum_id <- left_join(sum_id,G1, by="id")
# sum_id$G1_freq[is.na(sum_id$G1_freq)] <- 0
```

# Competitive interactions

```{r}
# total interaction instances (all species)
int_tot <- follows %>% 
  filter(initiator>=1) %>% 
  group_by(id) %>% summarise(int_tot=n())
# interaction instances with scarids only
int_scar <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","strp","rtp","ytp")) %>%
  group_by(id) %>% summarise(int_scar=n())
# interaction instances with conspecifics only
int_con <- follows %>% 
  filter(initiator>=1) %>% 
  filter(species==competitor_species) %>% 
  group_by(id) %>% summarise(int_con=n())
# interaction instances with heterospecific scarids
int_hetscar <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code=="stop" | competitor_code=="qup" | competitor_code=="rbp") %>% 
  filter(species!=competitor_species) %>% 
  group_by(id) %>% summarise(int_hetscar=n())
# interaction instances with damselfish
int_dam <- follows %>% 
  filter(initiator>=1) %>% 
  filter(competitor_code=="dam") %>% 
  group_by(id) %>% summarise(int_dam=n())

# join and add to sum_id dataframe
sum_id <- sum_id %>% 
  left_join(int_tot, by="id") %>% 
  left_join(int_scar, by="id") %>% 
  left_join(int_con, by="id") %>% 
  left_join(int_hetscar, by="id") %>% 
  left_join(int_dam, by="id")
sum_id[c("int_tot","int_scar","int_con","int_hetscar","int_dam")][is.na(sum_id[c("int_tot","int_scar","int_con","int_hetscar","int_dam")])] <- 0

# matrix of competitive interactions by species
stop <- follows %>% 
  filter(species_code=="stop") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(stop=n())
qup <- follows %>% 
  filter(species_code=="qup") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(qup=n())
rbp <- follows %>% 
  filter(species_code=="rbp") %>% 
  mutate(follows=n()) %>% 
  filter(competitor_code %in% c("stop","qup","rbp","prp","rtp","strp")) %>% 
  group_by(competitor_code,island) %>% summarize(rbp=n())

int_matrix <- full_join(stop, qup, by = c("competitor_code","island")) %>%
  full_join(rbp, by = c("competitor_code","island"))
```

# Summarizing data by site, species, and phase
Next I create a summary tables that averages variables within each site/species/phase, then combines into one table
```{r,warning=FALSE, message= FALSE}
# site averages
sum_site <- sum_id %>% 
  filter(length_cm >= 15 & length_cm <= 30) %>%
  group_by(site,island) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n),
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n),
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n),
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n),
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n),
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n)
    ) %>% 
  select(-n)

# site averages by species
sum_site_sp <- sum_id %>% 
  filter(length_cm >= 15 & length_cm <= 30) %>%
  group_by(site, island, species, species_code) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n), 
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n), 
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n), 
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n), 
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n), 
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n)
    ) %>% 
  select(-n)

# site averages by species and phase
sum_site_sp_ph <- sum_id %>% 
  filter(length_cm >= 15 & length_cm <= 30) %>%
  group_by(site, island, species, species_code, phase) %>% 
  summarize(
    n=n(),
    g_frac_mean=mean(g_frac),
    g_frac_se=sd(g_frac)/sqrt(n),
    br_mean=mean(br,na.rm=TRUE),
    br_se=sd(br,na.rm=TRUE)/sqrt(n),
    fr_mean=mean(fr),
    fr_se=sd(fr)/sqrt(n),
    for_dur_mean=mean(for_dur,na.rm=TRUE),
    for_dur_se=sd(for_dur,na.rm=TRUE)/sqrt(n),
    for_bites_mean=mean(for_bites,na.rm=TRUE),
    for_bites_se=sd(for_bites,na.rm=TRUE)/sqrt(n), 
    int_tot_mean=mean(int_tot),
    int_tot_se=sd(int_tot)/sqrt(n),
    int_scar_mean=mean(int_scar),
    int_scar_se=sd(int_scar)/sqrt(n),
    int_con_mean=mean(int_con),
    int_con_se=sd(int_con)/sqrt(n), 
    int_hetscar_mean=mean(int_hetscar),
    int_hetscar_se=sd(int_hetscar)/sqrt(n),
    int_dam_mean=mean(int_dam),
    int_dam_se=sd(int_dam)/sqrt(n),
    length_m = mean(length_cm),
    length_se = sd(length_cm)/sqrt(n)
    ) %>% 
  select(-n)
```

### 1_3 Join datasets, print .csv
```{r, echo=FALSE,message=FALSE,warning=FALSE}
# integrate all site-level variables
sum_site <- sum_site %>% 
  left_join(site, by=c("island","site"))
sum_id <- sum_id %>% 
  left_join(site, by=c("island","site")) %>%
  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
sum_site_sp <- sum_site_sp %>% 
  left_join(site, by=c("island","site")) %>%
  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)
sum_site_sp_ph <- sum_site_sp_ph %>% 
  left_join(site, by=c("island","site")) %>%
  left_join(sp_site, by=c("species","site")) # add conspecific scarid info (based on site/species)

write.csv(sum_site,file=here("data/sum_site.csv"))
write.csv(sum_id,file=here("data/sum_id.csv"))
write.csv(sum_site_sp,file=here("data/sum_site_sp.csv"))
write.csv(sum_site_sp_ph,file=here("data/sum_site_sp_ph.csv"))
```