### 2019-04-16 r/map.Rmd

---
title: "Site map"
author: "Molly Wilson"
date: "3/2/2019"
output: html_document
---

```{r}
library(ggplot2)
library(ggmap)

map.world <- map_data("world")

map.carib <- ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_cartesian(xlim = c(-100, -50),ylim = c(30,5)) +
  labs(x = "", y="",title = "")

ggplot() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_cartesian(xlim = c(-62, -60),ylim = c(18,17)) +
  labs(x = "", y="",title = "")

ggplot() +
  theme_bw() +
  geom_polygon(data = map.world, aes(x = long, y = lat, group = group))+
  coord_equal(xlim = c(-100, -60),ylim = c(27,10)) +
  labs(x = "", y="",title = "")
```