### 2019-05-10 r/5.8.19.Rmd
5/10

---
title: "GAMM exploration"
author: "Molly Wilson"
date: "5/8/2019"
output: html_document
---

Testing different GAMM iterations to confirm analyses.
Tasks at hand:
- site level analyses vs. individual fish level?
- include phases, or initial phase only?
- add site as random effect (in addition to island?)

Notes after exploration:
- Both site-level and id-level are 'possible', but still not sure which is most appropriate. Best models for id-level have many species interactions. I could split these models by species but obviously would lose a lot of R2.
- Will stick with initial phase only
- Site-level models for initial phase only cannot have 'method = "REML"'
- Site as random effect doesn't really change much in terms of best models or R2....

```{r, echo=F, message=F, warning=F}
library(here)
library(tidyverse)
library(mgcv)
library(MuMIn)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X) 
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% 
  filter(fr <= 3000)
sum_ssp <- read.csv(here("data","sum_ssp.csv"), stringsAsFactors=F) %>%
  select(-X) 
```

Assemble PC1 and PC2 variables
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, lc_cover, scar_den, scar_bm, carn30_bm)),
  center = T, scale = T)
pca_bent <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy)),
  center = T, scale = T)
pca_fish <- prcomp(na.omit(sum_site %>% 
              select(scar_bm, carn30_bm)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, shape = 'island', colour = 'island',
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(values = c("cadetblue3","cadetblue3","cadetblue3")) +
    theme_bw() +
    theme(legend.title = element_blank())
ggsave(here("figs/pca_pred.png"), width = 8, height = 5)

sum_site_pca <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)
sum_id_pca <- sum_id %>% 
  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site_pca,site,pc1_all,pc2_all), by = "site")

var <- get_pca_var
```

## Site - vs. ID - level GAMMs

ID - level, initial phase only
```{r, echo=F, message=F, warning=F}
m_data <- sum_id_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(fr ~ species_code + s(length_cm, k=5) + s(pc1_all, k=5) + s(pc2_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AIC(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr ~ species_code + s(length_cm, k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AIC(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr ~ species_code + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AIC(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr ~ species_code + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5) + s(pc2_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AIC(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr ~ species + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AIC(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr ~  s(length_cm, k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AIC(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr ~ s(length_cm, by = factor(species_code), k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AIC(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr ~ species + s(length_cm, by = factor(species_code), k=5) + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AIC(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr ~ species + s(length_cm, k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AIC(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr ~ species + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AIC(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr ~ s(length_cm, by = factor(species_code), k=5) + s(pc1_all, by = factor(species_code), k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AIC(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(fr ~ species + s(pc1_all, k=5), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
g12 <- c("g11", "Species + PC1", AIC(gamm11$lme), summary(gamm11$gam)$r.sq)

gtab_fr_id <- as.data.frame(rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AIC", "Adj. R2")) %>% arrange(AIC)

# plots from best models
plot(gamm3$gam,pages=1)
plot(gamm5$gam,pages=1)
```

Site - level GAMMs, initial phase only
```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(fr_mean ~ species_code + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)

gtab_fr_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)

# summary + plots from best models
summary(gamm12$gam)
plot(gamm12$gam,pages=1)
```

## Random effects: site vs. site + island

Site - level GAMM adding site as random effect
```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(fr_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

# gamm3 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
#            family = gaussian(link=identity),
#            random = list(island=~1, site=~1),
#            data = m_data
#            )
# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(fr_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

# gamm6 <- gamm(fr_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
#            family = gaussian(link=identity),
#            random = list(island=~1, site=~1),
#            data = m_data
#            )
# g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

# gamm8 <- gamm(fr_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
#            family = gaussian(link=identity),
#            random = list(island=~1, site=~1),
#            data = m_data
#            )
# g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(fr_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(fr_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(fr_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(fr_mean ~ species + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1, site=~1),
           data = m_data
           )
g12 <- c("g10", "Species + PC1", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gtab_fr_site_rd2 <- as.data.frame(rbind(g1,g2,g4,g5,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)

# plots from best models
plot(gamm10$gam,pages=1)
plot(gamm2$gam,pages=1)
```

# Feeding intensity
Site - level GAMMs, initial phase only
```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp" & phase == "i")

gamm1 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g1 <- c("g1", "Species + Size + PC1 + PC2", AICc(gamm1$lme), summary(gamm1$gam)$r.sq)

gamm2 <- gamm(for_bites_mean ~ species_code + s(length_mean, k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g2 <- c("g2", "Species + Size + PC1", AICc(gamm2$lme), summary(gamm2$gam)$r.sq)

# gamm3 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, by = factor(species_code), k=3),
#            family = gaussian(link=identity),
#            random = list(island=~1),
#            data = m_data
#            )
# g3 <- c("g3", "Species + Size~Species + PC1~Species + PC2~Species", AICc(gamm3$lme), summary(gamm3$gam)$r.sq)

gamm4 <- gamm(for_bites_mean ~ species_code + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3) + s(pc2_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g4 <- c("g4", "Species + Size~Species + PC1~Species + PC2", AICc(gamm4$lme), summary(gamm4$gam)$r.sq)

gamm5 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g5 <- c("g5", "Species + Size~Species + PC1~Species", AICc(gamm5$lme), summary(gamm5$gam)$r.sq)

gamm6 <- gamm(for_bites_mean ~ s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g6 <- c("g6", "Size + PC1~Species", AICc(gamm6$lme), summary(gamm6$gam)$r.sq)

gamm7 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g7 <- c("g7", "Size~Species + PC1", AICc(gamm7$lme), summary(gamm7$gam)$r.sq)

gamm8 <- gamm(for_bites_mean ~ species + s(length_mean, by = factor(species_code), k=3) + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g8 <- c("g8", "Species + Size~Species + PC1", AICc(gamm8$lme), summary(gamm8$gam)$r.sq)

gamm9 <- gamm(for_bites_mean ~ species + s(length_mean, k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g9 <- c("g9", "Species + Size + PC1~Species", AICc(gamm9$lme), summary(gamm9$gam)$r.sq)

gamm10 <- gamm(for_bites_mean ~ species + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g10 <- c("g10", "Species + PC1~Species", AICc(gamm10$lme), summary(gamm10$gam)$r.sq)

gamm11 <- gamm(for_bites_mean ~ s(length_mean, by = factor(species_code), k=3) + s(pc1_all, by = factor(species_code), k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g11 <- c("g11", "Size~Species + PC1~Species", AICc(gamm11$lme), summary(gamm11$gam)$r.sq)

gamm12 <- gamm(for_bites_mean ~ species_code + s(pc1_all, k=3), 
           family = gaussian(link=identity),
           random = list(island=~1),
           data = m_data
           )
g12 <- c("g12", "Species + PC1", AICc(gamm12$lme), summary(gamm12$gam)$r.sq)

gtab_fi_site <- as.data.frame(rbind(g1,g2,g4,g5,g6,g7,g8,g9,g10,g11,g12)) %>% setNames(c("g","Model", "AICc", "Adj. R2")) %>% arrange(AICc)

# plots from best models
plot(gamm12$gam,pages=1)
plot(gamm10$gam,pages=1)

summary(gamm12$gam)
```