### 2019-06-13 r/map_exp.Rmd
updating map/shapefiles

---
title: "Map"
author: "Molly Wilson"
date: "6/4/2019"
output: html_document
---

```{r}
library(tidyverse)
library(here)
library(sf)
library(ggmap)
# library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)

# site coordinates
site_data <- read.csv(here("data","map","site_coord.csv"))

# shapefiles for islands filtered from BES and ATG boundaries
bonaire <- st_read(here("data","map","BES_adm1.shp")) %>% 
  filter(NAME_1 == "Bonaire") %>%
  st_union() %>%
  st_sf() %>%
  mutate(island = "Bonaire")

atg <- st_read(here("data","map","ATG_adm1_gadm.shp")) 
antigua <- atg %>%
  filter(!NAME_1 == c("Barbuda","Redonda")) %>%
  st_union() %>%
  st_sf() %>%
  mutate(island = "Antigua")
barbuda <- atg %>% 
  filter(NAME_1 == "Barbuda") %>%
  st_union() %>%
  st_sf() %>%
  mutate(island = "Barbuda")

islands <- rbind(antigua, barbuda, bonaire) %>%
  mutate(lon=map_dbl(geometry, ~st_centroid(.x)[[1]]), 
         lat=map_dbl(geometry, ~st_centroid(.x)[[2]]))

# google API (not sure if I can register this permanently?)
register_google(key = "AIzaSyDvckfjy2riROUUuL9oDg7HBXRBd39X82A")
```

### Regional map

```{r}
world <- ne_countries(scale = "large", returnclass = "sf")
ggplot() +
  geom_sf(data = world, fill = "gray80", col = "gray80", size = .25) +
  geom_sf(data = islands, fill = "gray20", col = "gray20", size = .25) +
  #geom_text_repel(data = islands %>% filter(island %in% c("Antigua","Barbuda")), aes(x = lon, y = lat, label = island), nudge_x = 6, segment.size = .2) +
  #geom_text_repel(data = islands %>% filter(island == "Bonaire"), aes(x = lon, y = lat, label = island), nudge_x = -2, nudge_y = 2, segment.size = .2) +
  #annotate(geom = "text", x = -86.5, y = 26, label = "Gulf of Mexico", fontface = "italic", color = "grey80", size = 3) +
  annotate(geom = "text", x = -65, y = 29, label = "Atlantic Ocean", fontface = "italic", color = "grey80", size = 4) +
  annotate(geom = "text", x = -72.5, y = 15, label = "Caribbean Sea", fontface = "italic", color = "grey80", size = 4) +
  coord_sf(xlim = c(-86, -58), ylim = c(10, 33), expand = FALSE) +
  labs(x = "", y = "") +
  theme_bw()
ggsave(here("figs/map_region.png"))
```

Alternate (geocoding) option
```{r}
#world <- map_data("world") # df
ll.islands <- geocode(c("Antigua","Barbuda","Bonaire"))
islands.x <- ll.islands$lon
islands.y <- ll.islands$lat

ggplot() +
  theme_bw() +
  geom_sf(data = world, fill = "gray80", col = "gray80") +
  annotation_scale(location = "tr", width_hint = 0.25) +
  coord_sf(xlim = c(-90, -58), ylim = c(10, 33)) +
  geom_point(aes(x = islands.x, y = islands.y), color = "#f24db3", size = 5, shape = 1) +
  labs(x = "", y = "",title = "")

```

### Island-level details

```{r}
sites_anu <- site_data %>% filter(island == "Antigua")
ggplot() +
  geom_sf(data = antigua, fill = "slategray4", color = "slategray4") +
  geom_point(data = sites_anu, aes(x = lon, y = lat), color = "gray20", size = 2) +
  geom_text_repel(data = sites_anu,
                  aes(x = lon, y = lat, label = site),
                  segment.size = .2,
                  direction = "y",
                  box.padding = .75,
                  nudge_x = .18, nudge_y = .08,
                  size = 5) +
  coord_sf(xlim = c(-61.92, -61.52), ylim = c(16.98, 17.18)) +
  annotation_scale(location = "bl", width_hint = 0.25) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))

ggsave(here("figs/map_anu.png"))
```

```{r}
sites_brb <- site_data %>% filter(island == "Barbuda")
ggplot() +
  geom_sf(data = barbuda, fill = "slategray4", color = "slategray4") +
  geom_point(data = sites_brb, aes(x = lon, y = lat), color = "gray20", size = 2) +
  geom_text_repel(data = sites_brb,
                  aes(x = lon, y = lat, label = site),
                  segment.size = .2,
                  direction = "y",
                  box.padding = .75,
                  nudge_x = .18, nudge_y = .12,
                  size = 5) +
  coord_sf(xlim = c(-61.9, -61.44), ylim = c(17.51, 17.74)) +
  annotation_scale(location = "bl", width_hint = 0.25) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))

ggsave(here("figs/map_brb.png"))
```

```{r}
sites_bon <- site_data %>% filter(island == "Bonaire")
ggplot() +
  geom_sf(data = bonaire, fill = "slategray4", color = "slategray4") +
  geom_point(data = sites_bon, aes(x = lon, y = lat), color = "gray20", size = 2) +
  geom_text_repel(data = sites_bon,
                  aes(x = lon, y = lat, label = site),
                  segment.size = .2,
                  direction = "y",
                  box.padding = .75,
                  nudge_x = .22, nudge_y = .02,
                  size = 5) +
  coord_sf(xlim = c(-68.45, -67.85), ylim = c(12.02, 12.32)) +
  annotation_scale(location = "bl", width_hint = 0.25) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), panel.grid.major = element_line(color = "transparent"))

ggsave(here("figs/map_bon.png"))
```