### 2019-09-18 r/FID.Rmd

---
title: "Scarid Flight Response, Antigua"
author: "Molly Wilson"
date: "April 12, 2017"
output: html_document
---

```{r}
# prep
library(lazyeval)
library(here)
library(tidyverse)
library(reshape2)
```
```{r}
# import, format
fid <- read.csv(here("data","fid.csv"), stringsAsFactors=F) %>% 
  select(-c(FID.ft,FID.in,Bump.distance.ft,Bump.distance.in)) %>%
  janitor::clean_names()
```

# FID by species
```{r}
fid.sum <- fid %>%
  group_by(species) %>%
  summarize(n=n(),
            fid_mean = mean(fid_cm/100),
            fid_mean_se = sd((fid_cm/100))/sqrt(n),
            fid_max = max(fid_cm/100))

fid.ch2sum <- fid.sum %>%
  filter(species %in% c("Queen","Stoplight"))
```

# FID by spear, site
```{r}
lm.fid <-lm(data=fid, fid_cm ~ spear + site + rugosity + pre_approach_activity)
summary(lm.fid)

ggplot(fid, aes(x = pre_approach_activity,y=fid_cm)) + geom_point()
```

# bump distance by spear, site
```{r}
bump <- fid %>% filter(bump_distance_cm != 0)
lmBump <-lm(data=bump,bump_distance_cm ~ spear + site + species + rugosity + pre_approach_activity)
summary(lmBump)
aov(lmBump)

ggplot(bump, aes(x=spear,y=bump_distance_cm)) + geom_point(aes(color=species))
```