### 2019-09-18 r/Workflow.Rmd

---
title: "Workflow"
author: "Molly Wilson"
date: "5/8/2019"
output: html_document
---

Data processing and analaysis workflow for 'Scarid Behavior':

1. **Sites.rmd**: process site-level fish, benthic, and rugosity data.
*Exports*:
  - site.csv
  - sp_site.csv (fish species abundance/biomass data for later adding conspecific metrics)
  - fish.csv (processed fish data for later calculating grazing impacts of a given fish community)

2. **Follows.rmd**: process individual fish-level behavioral data, add site-level descriptors into individual-level and site-level summary datasets.
*Exports*:
  - sum_id.csv (indv. fish-level data, including site characteristics)
  - sum_site.csv (site level summary data)
  - sum_ss.csv (summary data by site and species)
  - sum_ssp.csv (summary data by site, species, and phase)

3. Exploration/analysis files:
  - **Data_prodding.rmd**: exploring relationships among variables, distributions, etc.
  - **Analysis.rmd**: initial boxplots/bargraphs looking at trends among islands/sites, testing ANOVAs/assumptions, predicted grazing impact, bite content calculations
  - **Models.rmd**: PCAs, LMEs, GAMMs (exploratory, no .csv exports)
  - **Ch2_outline_V3**: selected analyses for current manuscript outline